### 2013-06-12 paper/paper.tex
initial commit

\documentclass{sig-alternate}

\usepackage{listings}
\usepackage{hyperref}
\renewcommand{\topfraction}{0.9}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.84}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{ICER 2013}{'13 San Diego, CA USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{An Open Platform for Managing Short Programming Exercises}

\numberofauthors{2}

% \alignauthor
% Andrei Papancea\\
% Jaime Spacco\\
%        \affaddr{Department of Computer Science}\\
%        \affaddr{Knox College}\\
%        \affaddr{Galesburg, IL 61401}\\
%        \email{\{apapance,jspacco\}@knox.edu}
% % 2nd. author
% \alignauthor
% David Hovemeyer\\
%        \affaddr{Kinsley Engineering Center}\\
%        \affaddr{York College of Pennsylvania}\\
%        \affaddr{York, PA 17403}\\
%        \email{dhovemey@ycp.edu}

\author{
\alignauthor
Author Name\\
Author Name\\
       \affaddr{Department of Computer Science}\\
       \affaddr{Small College \#1}\\
       \affaddr{City, State zipcode}\\
       \email{\{author1,author2\}@school.edu}
% 2nd. author
\alignauthor
Author Name\\
       \affaddr{Computer Science Department}\\
       \affaddr{Small College \#2}\\
       \affaddr{City, State zipcode}\\
       \email{author@school.edu}
}

%\date{30 July 1999}

\maketitle
\begin{abstract}

We have built CloudCoder, an open platform for creating, assigning, and sharing
short programming exercises for a variety of languages (we currently
support C/C++, Java, Python and Ruby, and can add others).  Like other
systems, CloudCoder is a web service that lets
students write code directly in their web browser, click the
``submit'' button, and receive immediate feedback.  Unlike other
systems, which tend to be closed, or commercial, or both, CloudCoder
is a completely open platform.  The code for the system is
open-source, and exercises written for CloudCoder are shared to a
central repository under a permissive Creative Commons license
(BY-SA).  Finally, CloudCoder collects detailed data that faculty can use for
educational research.

In this work, we report on successful pilot studies of CloudCoder at several
institutions.  We also describe on-going work to create a ``hint'' system
for CloudCoder, similar to what one would expect from an Intelligent
Tutoring System (ITS).

% Students access CloudCoder exercises 
% through their web browser, where they can edit code

% Importance of understanding code more deeply.

% Identify ``knowledge components''

% Identify when and where to insert hints

% (Maybe) use Euclidean distance to compute closeness of two solutions

% To generate hints:  What hints to generate, when to generate them

\end{abstract}

% A category with the (minimum) three required fields
\category{K.3.2}{Computers and Education}{Computer and Information
  Sciences Education}[Computer Science Education]

\terms{Human Factors, Measurement}

\keywords{CS1, novice programmers, data collection, student behavior,
  intelligent tutoring system}

\section{Introduction}

Recent years have seen an explosion of online services for learning
to program:  Udacity \cite{udacity}, CodeAcademy \cite{codeacademy},
ProgZoo \cite{progzoo}, 
Codingbat \cite{codingbat}, Coursera \cite{coursera}, Turingscraft
\cite{turingscraft}, Problets \cite{Kumar::2005:GPA:1163405.1163408}, CS Circles \cite{Pritchard:2013:CCI:2445196.2445370} and
several others.  These systems are very powerful, and show great promise to
improve introductory computer science.  Given all of these
very good systems, one may ask why we thought it a good career move
to build yet another system.  There are three main reasons:

%\begin{itemize}

%\item 
First, we wanted a system that an instructor can easily integrate into an existing traditional course.
  Many massive open online courses (MOOCs), such those available through Udacity and
  Coursera, were not designed for a traditional classroom and cannot be easily incorporated into an existing course.  
  The infrastructure provided by a MOOC is geared towards the course
  taught through the MOOC, and is not really configurable or customizable.

 % systems, such as Udacity, Coursera, and CodeAcademy, are
 %  massive open online courses (MOOCs) that cannot be easily
 %  incorporated into a traditional classroom setting.  MOOCs can be wonderfully
 %  enriching for students who take these courses.  However, because
 %  MOOCs offer an entire
 %  course, whatever online programming infrastructure they provide is part of that
 %  course.

%\item 
Second, we wanted an open, customizable system that focuses on short programming exercises that students
  can solve in their browser, and receive immediate feedback.  There are excellent systems that help
  faculty manage longer programming assignments
  \cite{Edwards:2008:WAG:1384271.1384371}, or homework and review questions
  \cite{turingscraft, Kumar:2005:GPA:1163405.1163408}.  Codingbat
  \cite{codingbat} is a fantastic free service that provides short
  programming exercises, but it is not easily customizable and does not provide
  access to all of the submissions submitted by students.
  We see CloudCoder as a complementary component to these resources.

% Some existing systems, such as Problets \cite{Kumar:2005:GPA:1163405.1163408},  and Turingscraft
%   \cite{turingscraft}, provide homework exercises that can be incorporated
%   such as Codingbat \cite{codingbat},
%  , can be integrated into an existing course.
%   Codingbat, while free to use and fantastically useful, does not give faculty access to a very
%   rich data set

%\item
Finally, we wanted exercises that could be freely modified and
redistributed.  Much of the excellent content available through MOOCs,
or the increasing amount of digital content available through textbook publishers, 
cannot be modified or re-used outside of those contexts.  This can be a significant
impediment to researchers trying to study the difficulty level of
various exercises, or to build intelligent tutoring systems (ITSs)
that require a large bank of questions and their solutions.

%\end{itemize}

% There are two reasons.  First, many of
% these systems, such are Udacity, Coursera, and CodeAcademy, are
% massive open online courses (MOOCs).  MOOCs typically provide
% supporting infrastructure that helps students learn to code in their
% browsers, and the course material offered by a MOOC can be a wonderfully enriching
% experience for someone who enrolls in the course.  However, it is not
% clear how faculty who teach more traditional courses can
% integrate any of the infrastructure provided by a MOOC into
% their curriculum.  Second, 
% the systems that could be easily integrated into an existing course, such as
% Codingbat or Turingscraft, are
% are either commercial, or closed, or both.  To our knowledge, none of
% these systems readily support faculty who want to 
% modify the code, host their own installation, extract data about
% student behaviors beyond the basic correctness of their submissions,
% or redistribute the programming exercises and assignments.

We believe the academic community should have a fully open alternative
to the closed and/or commercial systems that have exploded onto the CS
Education scene.  Faculty can integrate CloudCoder into an existing course with minimal overhead,
can write their own content, use and modify content contributed by
others, customize the system, and collect data for educational studies.

\section{CloudCoder}
We have designed CloudCoder to be an open platform for managing short
programming exercises.  The CloudCoder platform has four design
goals:  To be an open-platform for faculty to assign short programming
exercises, to build a collection of redistributable programming
exercises, to collect data about student behaviors as they solve
exercises, and to explore ways of curating the collected data.

% \begin{itemize}
% \item {\em To provide a web-based platform for faculty to assign short
%   programming exercises.}

% There are platforms such as Web-CAT \cite{Edwards:2008:WAG:1384271.1384371},
% 
%  TODO Insert Marmoset reference!
%
 % and the many in-house systems developed at universities, for managing longer programming
 %  assignments, and there are commercial systems such as Turingscraft
 %  for assigning online homework \cite{turingscraft}.  However, there
  % are currently no open, flexible systems that serve a similar
  % purpose.  All of the code is open-source and available at Github:

% \item {\em To crowd-source the construction of a repository of programming
%   exercises.}  
% \item {\em To collect data about student work on short
%   programming exercises.}  There are many systems that collect data
%   about novice programmers \cite{Utting:2012:WDG:2361276.2361278,
%     Jadud:2006:MTE:1151588.1151600, Norris:2008:CCQ:1384271.1384284,
%     Spacco:2006:EMD:1140124.1140131}. % TODO: Cite French paper on Eclipse plugin
%   However, these systems typically target longer programming
%   assignments that students spend a week or two working on.
%   CloudCoder target very short exercises, and logs data at the
%   keystroke level.  We envision faculty using CloudCoder as a platform
%   for educational studies that require studying fine-grained trace data.
% \end{itemize}

\subsection{An Open Platform}

CloudCoder is an open platform for managing short programming
exercises.  Each exercise can be solved with one method or one short
program, and could take only a few minutes to solve, or could require
or an hour or two.  The system is not designed to support longer
programming assignments, such as those found at the {\em Nifty
  Assignments} session at SIGCSE.

Figure~\ref{screenshot} shows a screenshot of a CloudCoder exercise.
The editor is derived from the Ace editor \cite{ace}, and supports
syntax high-lighting for most programming languages, and
auto-indentation.  
A submit button runs the code against a series of
test cases, and feedback is provided directly in the browser.
CloudCoder captures anything the code prints to standard out or
standard err to help students with debugging.

\begin{figure*}[ht]
\centering
\includegraphics[width=180mm]{images/screenshot.png}
\caption{Screenshot of the main view in CloudCoder.}
\label{screenshot}
\end{figure*}

CloudCoder is available under the GNU Affero open-source license.
Binary distributions are available through our website (anonymized),
and the code is available on Github:

\vspace*{3mm} 
%\url{https://github.com/cloudcoderdotorg/CloudCoder}
(link anonymized)
\vspace*{3mm}

% The editor window

\subsection{Permissively-Licensed Exercises}

Writing good homework exercises is difficult and
time-consuming, and even the best textbooks are limited in the
number and quality of practice exercises they can provide, to say
nothing of a textbook's inability to provide timely feedback to
students.
Sharing exercises between faculty members is a much
more efficient way to increase the number and quality of exercises at
our disposal.

Faculty who use CloudCoder can write their own exercises using our in-browser
authoring tools, or can import exercises written by other faculty
from a central repository at:

\vspace*{3mm} 
% \url{https://cloudcoder.org/repo}
(link anonymized)
\vspace*{3mm}

Currently we have over 50 Java and 60 C/C++ exercises authored by 5
different users.  By default,
any new exercises written by faculty will be shared to the 
central repository under a permissive Creative Commons license (BY-SA)
that allows redistribution provided that the author is credited and
redistribution is under the same license.  
(Note that the repository shows the exercise description, but not the
test cases or the solution, so an unscrupulous student cannot dig through the
repository for solutions.  Importing exercises requires an account at
the repository, which can be used to discourage this sort of
undesirable behavior.).

CloudCoder supports two types of exercises:  {\em method} or {\em function-based}
exercises, and {\em whole-program} exercises.  Function based exercises look
a lot like Codingbat exercises:  Students are given a description of
what the method should do, and a couple of example input/output pairs
similar to a unit test.  Writing a new function-based exercise
requires a description, some ``starter code'' (typically the method
header, although exercises can provide as much or as little code as
desired), and a series of inputs and their corresponding outputs.
Test-cases can be marked as ``secret'' so that students don't see their
outputs to disincentivize students from writing a special case to
recognize each input.

On the other hand, whole-program exercises
have a description, but read inputs from
standard-in and write to standard-out.  CloudCoder uses a
simplified regular expression language to check student outputs for
correctness.

\subsection{Building and Testing Submissions}

CloudCoder currently supports exercises for C/C++, Java, Python and
Ruby.  Building and testing takes place in a separate process which
can be run on a separate machine for extra security, or on multiple
machines for scalability.

We run Java code relatively securely under a Security Manager, and 
use Jython and JRuby to compile Python and Ruby to Java
bytecode, which we also run in a Security-Managed sandbox.  We
currently support C/C++ less securely, although 
recent work by Malan \cite{Malan:2013:CSS:2445196.2445242} on a
``sandbox'' for running untrusted student code can be leveraged to improve our security.

\subsection{Data Collection and Curation}

CloudCoder logs keystrokes in the online editor window, and 
stores all submissions and test outcomes in a databases.
We hope that the combination of exercises and data will give faculty
insights into the difficulty level of different exercises, keep a
record of common error patterns, and enable the development of
innovative data analyses.

One question is how to manage the data collected.  
Recent work on web-scale data collection by the BlueJ team
\cite{Utting:2012:WDG:2361276.2361278} lays out a number of important
questions about programming data, such as who should have access to the data, and what
mechanism should be used to access the data.  Sanders
et. al. \cite{Sanders:2008:DSE:1404520.1404534} reviewed a number of
the challenges involved in sharing empirical education data.

An alternative repository for the type of data collected by CloudCoder
is the Pittsburgh Science and Learning Center (PSLC) Datashop, which
is a repository of data collected by by intelligent tutoring
systems (ITSs).  By its nature, ITS data is much simpler than program
snapshot data, so snapshots cannot be added without some changes.
The Datashop is interested in storing program
snapshot data to make it easier for researchers in the ITS, educational data
mining, and CS-ed communities to work on programming data.

Anonymized data that has been cleared by the proper Institutional
Review Boards is contributed to the Datashop, which
makes the data available to other researchers under whatever
restrictions were put in place by the overseeing IRBs.  The Datashop also
provides a suite of analysis and visualization tools for better making
sense of the ITS data.

We are working with members of the Datashop to establish standards for
program snapshot datasets, and to begin developing tools and
techniques that can help analyze this very different dataset.  This is
a long-term project that is in its infancy.

% The Datashop stores traces from ITSs, which are typically made up of
% multiple choice questions, and merely need to store a student's
% answer, and whether they asked for a hint.

% We have been in contact with the the Pittsburgh Science and Learning
% Center (PSLC) Datashop.  The Datashop is a repository for

% The data collected by CloudCoder is simpler than the data collected by
% the BlueJ team.

%Standard storage formats, visualization tools, standard analyses.

\subsection{Hosting CloudCoder}
We recognize that ease of installation is a critical factor for
adoption of a new pedagogical
technology for busy
faculty who may only have a few minutes to evaluate a new system.  We
are committed to making the installation process as simple and painless as 
possible.

Thus far each institution that has adopted CloudCoder has
provided a Linux server that we configured for them.  This has made
adoption very easy, but is unlikely to scale up as we acquire more
users.  We have asked Amazon for Amazon Web Services credits to
provide CloudCoder ``in the cloud'' to adopting institutions, and we
hope to add support for Google App Engine this summer.  Ease of
adoption is critical for CloudCoder.

\subsection{Pilot Studies}

We piloted CloudCoder at several institutions during the 2012-2013
academic year.  Table~\ref{tab:courses} shows a breakdown of the
courses.  We tried to use CloudCoder at Auckland in Fall 2012 for a
course with 700 students, but the free-tier of Amazon EC2 did not
scale up well enough, so we only recorded partial data.  We have made
significant improvements since then and plan to deploy CloudCoder for
the same large course in Fall 2013.

\begin{table*}
\centering
\begin{tabular}{| l | l | l | l |}
\hline
Institution & Course & Language & Enrollment\\
\hline
Canisius College & CS-1 & C/C++ & 30\\
Small College \#2 & CS-1 & C/C++ & 120\\
Small College \#1 & CS-1 & Java & 23\\
University of Auckland & Programming for Engineers & C & 30\\
\hline
\end{tabular}
\caption{CloudCoder pilot studies.}
\label{tab:courses}
\end{table*}

In the Winter 2013 term at Small College \#1, students taking CS-1 used {\tt CloudCoder}
%
% TODO: \cite{Hovemeyer:2013:CBC:2445196.2445451}.  
% 
We collected over 6,000 submissions
for 32 exercises by 23 students.  We expect to have a much larger
dataset from 120 students at 
Small College \#2
%York College 
once the Spring 2013 semester ends.  Our analysis of these datasets is
very preliminary at this point, but work is ongoing.

\section{Discussion}

We could use community input into the CloudCoder project.  Specifically:

\begin{itemize}
\item Who would be interested in adopting this system?  If you would
  consider trying out the system if we added a particular feature,
  what would that feature be?  We would like
  to expand our pilot group.
\item What studies could or should be done with this platform?  And
  would you like to be involved with these studies?
\item What studies can be done with the existing data we have
  collected?  We do not yet have well-developed tools for data
  analysis, but one data set has been anonymized and can be
  distributed to other researchers interested in our data.
\end{itemize}

\section{Related Work}

Jadud's pioneering early work on novice compilation behavior
\cite{Jadud:2006:MTE:1151588.1151600}, along with follow-up work by
others \cite{Norris:2008:CCQ:1384271.1384284} provides a starting point for
work with program snapshots.

Denny et. al. have studied novice syntax errors
\cite{Denny:2012:SEE:2325296.2325318} using data collected from the CodeWrite system.
CodeWrite is similar to CloudCoder, except that students create
exercises and test cases for each other to solve, and are evaluated on
the quality of their practice exercises.

Lane et. al. describe Intention-Based Scoring (IBS) \cite{Lane:2005:ISA:1047344.1047471},
an approach to building an ITS for programming whereby human
judges to score or code {\em charettes}, or assignments students work on in a
timed, closed-lab environment.  IBS derives from prior work on
identifying ``bugs'' in online protocols, and as such attempts to
identify places where students have a ``bug'' in their development
process.  This approach, while labor-intensive, shows some promise in
developing an ITS for programming assignments.

Zingaro et. al. built the Python Classroom Response System
\cite{Zingaro:2013:FCP:2445196.2445369}, a system similar to
CloudCoder, except that it focuses on supporting in-class programming
questions for a class that uses a {\em Peer Instruction} pedagogical approach.

Pritchard et. al. present CS Circles
\cite{Pritchard:2013:CCI:2445196.2445370}, an interactive on-line
Python curriculum that embeds short questions, some of which involve
programming, directly into the text.  This sort of interactive text
shows great promise for developing open, interactive courses.

The BlueJ team presents a compelling case for web-scale data
collection \cite{Utting:2012:WDG:2361276.2361278}.  Their system is
more ambitious than CloudCoder, but the data collected by CloudCoder
should be complementary.

\section{Future Work: Towards an Intelligent Tutoring System}

One of the long-term goals for CloudCoder is to use the data to
develop an Intelligent Tutoring System for programming assignments.
A web-based programming ITS could be used to support high school
AP courses in school districts that may lack computer science
teachers, or to help improve introductory CS courses by reducing
student frustration.

To develop an ITS, we need a large supply of exercises; we hope
CloudCoder's permissible licensing will help on this front.  We also
need to provide students with ``hints'' when they are stuck.  To
provide hints, we can either write the hints ourselves, which is
labor-intensive, or generate them automatically by mining them out of
the dataset of prior submissions.  Obviously this second approach is
a much preferred method of creating hints.  However, identifying hints
automatically means that we need to match a student's current program
state with the most similar program state we've seen previously, 
figure out what successful next step the previous student took, and
then summarize this information into a hint.  This is a very
challenging problem.

There are currently two promising lines of research
trying to address the issue of identifying the most similar previous submissions.  Rivers and Koedinger \cite{rivers-its2012} have developed a
system that computes a {\em canonicalized} version of each submission
by compiling to an Abstract Syntax Tree (AST) and then applying a
series of compiler transformations.  This work is ongoing, but
preliminary results are encouraging.
On the other hand, Jin et. al. \cite{Jin:2012:PRA:2345840.2345889, jin-kdd2011} perform
an inter-procedural data flow analysis that tracks the values stored
in different variables in an effort to identify variables that are
used in the same way in each submission.  They also show promising preliminary
results for their approach.

\subsection{Key Features:  Mining Hints from the AST}

Our approach, which is very preliminary, ongoing research,  is also based on an analysis of
the AST.
Our hypothesis is that many novice-level exercises
will have solutions that require certain structural
{\em key features} in the code that we can detect automatically by
comparing the AST of the student's code with the ASTs of known correct
solutions.

For example, Figures~\ref{fig:counta} and \ref{fig:countab} show two
CloudCoder exercises that were assigned during the Winter 2013 term in
CS-1 at 
Small College \#1
%Knox College.

\begin{figure}[h]
{\tt Write a static method called countA that takes a String as a
parameter and returns the number of times that 'a' occurs in the String. }
\caption{CountA:  An example CloudCoder exercise.}
\label{fig:counta}
\end{figure}

\begin{figure}[h]
{\tt Write a static method called countAB that takes a String as a
parameter and returns the number of times that ab occurs (i.e. an 'a'
immediately followed by 'b'). }
\caption{CountAB: Another example CloudCoder exercise.}
\label{fig:countab}
\end{figure}

First let us consider the {\tt CountA} exercise from
Figure~\ref{fig:counta}.  As is typical in computer science, students have a lot of flexibility
in how they solve this problem:  with a for loop or with a while loop, using the {\tt charAt()}
method or the {\tt substring()} method or other String methods, and so on.  However, in
practice, exercises are assigned in a context with other
exercises, in-class examples, and course topics, which tends to limit
the scope of possible student solutions, and to push students towards
certain solutions and away from others.  For example, no student solved this
exercise with a while loop, and all but one used {\tt charAt()}
rather than other String operations.  In fact, most correct
solutions looked something like Figure~\ref{fig:counta-solution}.

\begin{figure}[h]
\begin{lstlisting}
public static int countAB(String str) {
  int count=0;
  for (int i=0; i<str.length();i++) {
    if (str.charAt(i)=='a') {
      count++;
    }
  }
  return count;
}
\end{lstlisting}
\caption{Typical solution for Exercise CountA depicted in Figure~\ref{fig:counta}.}
\label{fig:counta-solution}
\end{figure}

\begin{figure}
\includegraphics[width=90mm]{images/countA.pdf}
\caption{Slightly simplified AST of the code in Figure~\ref{fig:counta-solution}}
\label{fig:counta-ast}
\end{figure}

Compiling the method from Figure~\ref{fig:counta-solution} to an
Abstract Syntax Tree (AST) produces a tree that looks something like
Figure~\ref{fig:counta-ast} (the AST has been simplified slightly by
collapsing nodes that are unimportant for this example).  We can identify
several features of the AST that must be present to solve the
exercise, such as a for loop, an if-statement inside the loop, and
the use of a String method, typically {\tt charAt()}, to examine
characters inside the loop.  We can also ensure that the components of
the for loop are correct, i.e. that the loop counter starts at 0, that
the loop goes up to the length of the String, and so on.  Assuming
that student code compiles, all of these necessary features can be
extracted from the AST.

Now let's consider the {\tt CountAB} exercise depicted in
Figure~\ref{fig:countab}.  Again, there were many correct solutions;
however, many correct solutions looked something like
Figure~\ref{fig:countab-solution}.

\begin{figure}[h]
\begin{lstlisting}
public static int countAB(String str) {
  int count=0;
  for (int i=0; i<str.length()-1;i++) {
    if (str.charAt(i)=='a' && 
        str.charAt(i+1)=='b') {
      count++;
    }
  }
  return count;
}
\end{lstlisting}
\caption{Typical solution for Exercise countAB depicted in Figure~\ref{fig:countab}.}
\label{fig:countab-solution}
\end{figure}

Compiling the method in Figure~\ref{fig:countab-solution} into an
abstract syntax tree (AST) produces a tree that looks like
Figure~\ref{fig:countab-ast} (this result has been slightly simplified
from the full output from the Eclipse JDT compiler).

\begin{figure}[h]
\includegraphics[width=90mm]{images/forloop-highlight.png}
\caption{Slightly simplified AST of the code in Figure~\ref{fig:countab-solution}}
\label{fig:countab-ast}
\end{figure}

Note the two circled sub-trees, or {\em features}, in the AST depicted
in Figure~\ref{fig:countab-ast}.  Feature \#1, the {\em
  length-1} feature, shows the loop iterating
until the length of the String minus one.  The for loop requires a
stop condition that is different
than the loop stop condition used to solve the {\tt CountA} exercise from
Figure~\ref{fig:counta}.  Crucially, once we know students are
using a for loop to solve this exercise (which we can also
extract from the AST), {\em we know exactly where in the AST to
  look for the length-1 feature}.  If a submission has all the
features except the {\em length-1} feature, we have a pretty good idea
that the student hasn't figured out to change the stop condition of the
for loop.  We could use the lack of features present in correct
solutions to give us an idea of what things the student is missing.

\subsection{Limits of Key Features Approach}

Our approach of identifying ``key features'' has a number of
weaknesses and limitations.  First, like other ITS hint-systems, it
requires a reliable system for identifying the nearest previous successful
submission to a student's current program state.  Work in this area is
ongoing \cite{rivers-its2012,Jin:2012:PRA:2345840.2345889}.

Second, features are based purely on the structure of the
code, and are treated independently of one another, which is a
simplification that throws out a lot of relevant contextual information.  For
example, a student submission may have two for loops, one of which has
an if statement nested inside it; our approach cannot determine which
loop has the if statement inside it.

Third, we do not perform any data-flow analysis (at least not yet), so
we cannot identify if particular values are stored in variables.

Next, we require compilable code, which can be a problem for the many
students who struggle to get their code to compile
\cite{Denny:2012:SEE:2325296.2325318}.

Finally, while we focused mainly on exercises that required for loops
to solve, we also assigned a number of exercises that
use if statements and no loops.  For these exercises, we can figure
out when students don't have 
enough structure, such as when they don't have nested if statements
for an exercise where most people who solves the exercise used nested if statements.  However,
we cannot yet easily determine when students have the wrong boundary
conditions in an if statement.

Much more work is needed in this area.  We hope that 
CloudCoder can help out by providing a rich data set for post-hoc
studies, as well as a platform for testing out a hint-system.

\bibliographystyle{abbrv}
\bibliography{paper}

%\appendix
%Appendix A
\section{Acknowledgements}

Anonymized.

\end{document}

### 2013-06-12 analysis.r

@@ -1,0 +1,78 @@
+features=c('f1', 'f2', 'f3', 'f4', 'f5', 'f6', 'f7', 'f8', 'f9', 'f10', 'f11', 'f12', 'f13', 'f14', 'f15', 'f16', 'f17', 'f18', 'f19', 'f20', 'f21', 'f22', 'f23', 'f24', 'f25', 'f26', 'f27', 'f28', 'f29', 'f30', 'f31', 'f32', 'f33', 'f34', 'f35', 'f36', 'f37', 'f38', 'f39', 'f40', 'f41', 'f42', 'f43', 'f44', 'f45', 'f46', 'f47', 'f48', 'f49', 'f50', 'f51', 'f52', 'f53', 'f54', 'f55')
+featureDeltas=c('fd1', 'fd2', 'fd3', 'fd4', 'fd5', 'fd6', 'fd7', 'fd8', 'fd9', 'fd10', 'fd11', 'fd12', 'fd13', 'fd14', 'fd15', 'fd16', 'fd17', 'fd18', 'fd19', 'fd20', 'fd21', 'fd22', 'fd23', 'fd24', 'fd25', 'fd26', 'fd27', 'fd28', 'fd29', 'fd30', 'fd31', 'fd32', 'fd33', 'fd34', 'fd35', 'fd36', 'fd37', 'fd38', 'fd39', 'fd40', 'fd41', 'fd42', 'fd43', 'fd44', 'fd45', 'fd46', 'fd47', 'fd48', 'fd49', 'fd50', 'fd51', 'fd52', 'fd53', 'fd54', 'fd55')
+
+#
+# This gives the best scores, with the features merged back in
+#
+bestscores = function(mysubs) {
+  agg=aggregate(cbind(correcttests, t) ~ user*problem, data=subs, FUN=max)
+  merge(agg, subs, by=c('user','problem','t','correcttests'))
+}
+
+perfectscores = function(mysubs) {
+  mysubs=subset(mysubs, correcttests/numtests == 1)
+  return(bestscores(mysubs))
+}
+
+bools = function(mysubs) {
+  lambda=function(x) {
+    as.numeric(x != 0)
+  }
+  for(i in 7:62) {
+    mysubs[,i]=lambda(mysubs[,i])
+  }
+  return(mysubs)
+}
+
+#
+# For the love of the R gods,
+# I have no idea why this works!~
+# Why does this work, but not the
+# 5 variations I tried previously?
[…]

### 2013-06-12 crapfilter.py

@@ -1,0 +1,30 @@
+#!/usr/bin/env python
+
+from funs import *
+
+def flatten(lst, res=[]):
+    for x in lst:
+        if isinstance(x, list):
+            flatten(x, res)
+        else:
+            res.append(x)
+    return res
+
+keep=flatten([1,range(5,60)])
+#print flatten(keep)
+
+for line in open('newfeatures.csv'):
+    sub=line.rstrip().split()
+    if user(sub)=='user':
+        tmp=[]
+        for i in keep:
+            tmp.append(sub[i])
+        print '\t'.join(tmp)
+        continue
+    res=[]
+    if numtests(sub)>0 and numtests(sub)==correcttests(sub):
+        for i in keep:
+            res.append(sub[i])
+        print '\t'.join(res)
+
+

### 2013-06-12 deltas.py

@@ -1,0 +1,141 @@
+#!/usr/bin/env python
+
+import sys
+from funs import *
+
+def addIfNotExists(map, key, type='list'):
+    if not map.has_key(key):
+        if type=='list':
+            map[key]=[]
+        elif type=='map':
+            map[key]={}
+        else:
+            raise Exception('Unknow type to put into a map: %s' % type)
+
+class SubMap:
+    def __init__(self):
+        self.problems={}
+        self.users={}
+        self.allsubs=[]
+    def add(self, sub):
+        p=prob(sub)
+        u=user(sub)
+        # add to list of all subs
+        self.allsubs.append(sub)
+        # map: problem => user => list of subs
+        addIfNotExists(self.problems, p, 'map')
+        addIfNotExists(self.problems[p], u, 'list')
+        self.problems[p][u].append(sub)
+        # map: user => problem => list of subs
+        addIfNotExists(self.users, u, 'map')
+        addIfNotExists(self.users[u], p, 'list')
+        self.users[u][p].append(sub)
+    def computeDeltas(self,binary=False):
+        for (u,probs) in self.users.iteritems():
+            for (p,subs) in probs.iteritems():
+                computeDeltas(subs,binary)
+    def replaceUP(self, u, p, sub):
+        self.users[u][p]=sub
+        self.problems[p][u]=sub
+    def getSubsUP(self, u, p):
+        return self.user[u][p]
+    def __len__(self):
+        return len(self.allsubs)
+    def getUsers(self):
+        return sorted(self.users.keys())
+    def getProblems(self):
+        return sorted(self.problems.keys())
+    def __str__(self):
+        res=''
+        i=0
+        for (u,probs) in self.users.iteritems():
+            for (p,subs) in probs.iteritems():
+                for s in subs:
+                    res += '\t'.join([str(x) for x in s])
+                    res+='\n'
+        return res
+
+def parse(filename):
+    meta=SubMap()
+    for line in open(filename):
+        line=line.rstrip()
+        # skip header line
+        if line.startswith('user'):
+            continue
+        sub=line.split()
+        meta.add(sub)
+    return meta
+
+def addFields(sub, testdelta, fixcomp, featureDeltas=[0]*55):
+    sub.append(testdelta)
+    sub.append(fixcomp)
[…]

### 2013-06-12 dist.py

@@ -1,0 +1,56 @@
+#!/usr/bin/env python
+
+from funs import *
+
+def dist(f1, f2):
+    dsquare=0
+    if len(f1) != len(f2):
+        raise Exception("Feature lists must have the same length to compute the distance between them")
+    for i in range(len(f1)):
+        x1=int(f1[i])
+        x2=int(f2[i])
+        dsquare+=(x2-x1)*(x2-x1)
+    return dsquare
+        
+
+def compare(sub1, sub2, numFeatures=55):
+    i1=f(1, sub1)
+    i2=f(2, sub2)
+    return dist(sub1[i1:i1+numFeatures+1], sub2[i2:i2+numFeatures+1])
+
+def getPerfect(subs, problem):
+    perfect={}
+    for s in subs:
+        if prob(s)==problem and numtests(s) > 0 and correcttests(s) == numtests(s):
+            perfect[subnum(s)]=s
+    return perfect
+
+def getClosestMap(subs, problem):
+    closest={}
+    perfectMap=getPerfect(subs, problem)
+    #print len(perfectMap)
+    perfect=perfectMap.values()
+    for s in subs:
+        num=subnum(s)
+        if prob(s) != problem:
+            continue
+        if perfectMap.has_key(num):
+            continue
+        
+        d=compare(s, perfect[0])
+        closest[num]=(d, perfect[0])
+        for p in perfect:
+            c=compare(s, p)
+            if c<closest[num][0]:
+                closest[num]=(c, p)
+    return closest
+
+def main():
+    subs=[line.rstrip().split() for line in open('newfeatures.csv')]
+    #print len(subs)
+    map=getClosestMap(subs, '16')
+    print len(map)
+    for (s
+
+if __name__=='__main__':
+    main()

### 2013-06-12 extra.py

@@ -1,0 +1,7 @@
+#(user, problem, submission, numtests, correcttests, f1, f2, f3, f4, f5, f6, f7, f8, f9, f10, f11, f12, f13, f14, f15, f16, f17, f18, f19, f20, f21, f22, f23, f24, f25, f26, f27, f28, f29, f30, f31, f32, f33, f34, f35, f36, f37, f38, f39, f40, f41, f42, f43, f44, f45, f46, f47, f48, f49, f50, f51, f52, f53, f54, f55, t) = entry
+
+ALLFIELDS=['user', 'problem', 'submission', 'numtests', 'correcttests', 'f1', 'f2', 'f3', 'f4', 'f5', 'f6', 'f7', 'f8', 'f9', 'f10', 'f11', 'f12', 'f13', 'f14', 'f15', 'f16', 'f17', 'f18', 'f19', 'f20', 'f21', 'f22', 'f23', 'f24', 'f25', 'f26', 'f27', 'f28', 'f29', 'f30', 'f31', 'f32', 'f33', 'f34', 'f35', 'f36', 'f37', 'f38', 'f39', 'f40', 'f41', 'f42', 'f43', 'f44', 'f45', 'f46', 'f47', 'f48', 'f49', 'f50', 'f51', 'f52', 'f53', 'f54', 'f55', 't']
+
+FEATURES=['f1', 'f2', 'f3', 'f4', 'f5', 'f6', 'f7', 'f8', 'f9', 'f10', 'f11', 'f12', 'f13', 'f14', 'f15', 'f16', 'f17', 'f18', 'f19', 'f20', 'f21', 'f22', 'f23', 'f24', 'f25', 'f26', 'f27', 'f28', 'f29', 'f30', 'f31', 'f32', 'f33', 'f34', 'f35', 'f36', 'f37', 'f38', 'f39', 'f40', 'f41', 'f42', 'f43', 'f44', 'f45', 'f46', 'f47', 'f48', 'f49', 'f50', 'f51', 'f52', 'f53', 'f54', 'f55']
+
+features=range(1,56)

### 2013-06-12 foo.r

@@ -1,0 +1,38 @@
+#
+# TODO
+# 
+# How to handle stats for the same student across multiple exericses
+#
+
+# check num rows
+#nrow(subs)
+# check num cols
+#ncol(subs)
+# 36 total
+
+# remove data
+#
+#subs2=subset(subs, numtests>0)
+
+#
+# Find the max score
+# also the latest max score submission
+# 
+# I think that the cbind() here is telling us to max two cols
+# in the order they are given
+#
+# Note that this doesn't include the features, sadly
+#
+#agg=aggregate(cbind(correcttests, t) ~ user*problem, data=subs, FUN=max)
+#m1=merge(agg, subs, by=c('user','problem','t','correcttests'))
+
+#
+# This defaults into scientific notation because t is so large
+# We can mess with the format.
+#
+#format(colMeans(best), nsmall=2, scientific=FALSE)
+
+#
+# Can select a range of columns like this:
+#
+#colMeans(best[7:36])

### 2013-06-12 funs.py

@@ -1,0 +1,20 @@
+def subnum(sub):
+    return sub[2]
+def user(sub):
+    return sub[0]
+def prob(sub):
+    return sub[1]
+def correcttests(sub):
+    return int(sub[4])
+def numtests(sub):
+    return int(sub[3])
+def f(num, sub):
+    return int(sub[num+4])
+def fb(num, sub):
+    if sub[num+4]>0:
+        return 1
+    return 0
+def t(sub):
+    return sub[60]
+def iscompile(sub):
+    return numtests(sub) > 0

### 2013-06-12 gdoc.py

@@ -1,0 +1,18 @@
+#!/usr/bin/env python
+
+for line in open('uniq-subs.txt'):
+    (user_id, problem_id) = line.rstrip().split()
+    if user_id == 'user_id':
+        continue
+    required=''
+    if int(problem_id) in (14, 15, 18, 23, 24,
+                           25, 26, 16, 37, 38,
+                           39, 40, 41, 33, 45,
+                           42, 44, 46, 43, 48,
+                           49, 50, 51, 52, 30,
+                           28, 31, 32):
+        required='REQ'
+    if int(user_id) <= 6 or int(user_id) == 82:
+        required='INS'
+    print "%s\t%s\t%s\thttps://cs.knox.edu/apapance/analyzer/?u=%s&p=%s" % (user_id, problem_id, required, user_id, problem_id)
+

### 2013-06-12 go.py

@@ -1,0 +1,11 @@
+#!/usr/bin/env python
+
+features=[]
+for x in range(1,56):
+    features.append("'f%d'"%x)
+print 'features=c(' + ', '.join(features) + ')'
+
+featureDeltas=[]
+for x in range(1,56):
+    featureDeltas.append("'fd%d'"%x)
+print 'featureDeltas=c(' + ', '.join(featureDeltas) + ')'

### 2013-06-18 paper/paper.tex
restore author info
comment out sec 5, fix citation

@@
 
-% \alignauthor
-% Andrei Papancea\\
-% Jaime Spacco\\
-%        \affaddr{Department of Computer Science}\\
-%        \affaddr{Knox College}\\
-%        \affaddr{Galesburg, IL 61401}\\
-%        \email{\{apapance,jspacco\}@knox.edu}
-% % 2nd. author
-% \alignauthor
-% David Hovemeyer\\
-%        \affaddr{Kinsley Engineering Center}\\
-%        \affaddr{York College of Pennsylvania}\\
-%        \affaddr{York, PA 17403}\\
-%        \email{dhovemey@ycp.edu}
-
 \author{
 \alignauthor
-Author Name\\
-Author Name\\
+Andrei Papancea\\
+Jaime Spacco\\
        \affaddr{Department of Computer Science}\\
-       \affaddr{Small College \#1}\\
-       \affaddr{City, State zipcode}\\
-       \email{\{author1,author2\}@school.edu}
+       \affaddr{Knox College}\\
+       \affaddr{Galesburg, IL 61401}\\
+       \email{\{apapance,jspacco\}@knox.edu}
 % 2nd. author
 \alignauthor
-Author Name\\
-       \affaddr{Computer Science Department}\\
-       \affaddr{Small College \#2}\\
-       \affaddr{City, State zipcode}\\
-       \email{author@school.edu}
+David Hovemeyer\\
+       \affaddr{Kinsley Engineering Center}\\
+       \affaddr{York College of Pennsylvania}\\
+       \affaddr{York, PA 17403}\\
+       \email{dhovemey@ycp.edu}
 }
+
+%\author{
+%\alignauthor
+%Author Name\\
+%Author Name\\
+%       \affaddr{Department of Computer Science}\\
+%       \affaddr{Small College \#1}\\
+%       \affaddr{City, State zipcode}\\
+%       \email{\{author1,author2\}@school.edu}
+%% 2nd. author
+%\alignauthor
+%Author Name\\
+%       \affaddr{Computer Science Department}\\
+%       \affaddr{Small College \#2}\\
+%       \affaddr{City, State zipcode}\\
+%       \email{author@school.edu}
+%}
 
@@
 Codingbat \cite{codingbat}, Coursera \cite{coursera}, Turingscraft
-\cite{turingscraft}, Problets \cite{Kumar::2005:GPA:1163405.1163408}, CS Circles \cite{Pritchard:2013:CCI:2445196.2445370} and
+\cite{turingscraft}, Problets \cite{Kumar:2005:GPA:1163405.1163408}, CS Circles \cite{Pritchard:2013:CCI:2445196.2445370} and
 several others.  These systems are very powerful, and show great promise to
@@
 \centering
-\includegraphics[width=180mm]{images/screenshot.png}
+\includegraphics[width=180mm]{images/screenshot}
 \caption{Screenshot of the main view in CloudCoder.}
@@
 Canisius College & CS-1 & C/C++ & 30\\
-Small College \#2 & CS-1 & C/C++ & 120\\
-Small College \#1 & CS-1 & Java & 23\\
+York College & CS-1 & C/C++ & 120\\
+Knox College \#1 & CS-1 & Java & 23\\
 University of Auckland & Programming for Engineers & C & 30\\
@@
 
-\section{Future Work: Towards an Intelligent Tutoring System}
-
-One of the long-term goals for CloudCoder is to use the data to
-develop an Intelligent Tutoring System for programming assignments.
-A web-based programming ITS could be used to support high school
-AP courses in school districts that may lack computer science
-teachers, or to help improve introductory CS courses by reducing
-student frustration.
-
-To develop an ITS, we need a large supply of exercises; we hope
-CloudCoder's permissible licensing will help on this front.  We also
-need to provide students with ``hints'' when they are stuck.  To
-provide hints, we can either write the hints ourselves, which is
-labor-intensive, or generate them automatically by mining them out of
-the dataset of prior submissions.  Obviously this second approach is
-a much preferred method of creating hints.  However, identifying hints
-automatically means that we need to match a student's current program
-state with the most similar program state we've seen previously, 
-figure out what successful next step the previous student took, and
-then summarize this information into a hint.  This is a very
-challenging problem.
-
-There are currently two promising lines of research
-trying to address the issue of identifying the most similar previous submissions.  Rivers and Koedinger \cite{rivers-its2012} have developed a
-system that computes a {\em canonicalized} version of each submission
-by compiling to an Abstract Syntax Tree (AST) and then applying a
-series of compiler transformations.  This work is ongoing, but
-preliminary results are encouraging.
-On the other hand, Jin et. al. \cite{Jin:2012:PRA:2345840.2345889, jin-kdd2011} perform
-an inter-procedural data flow analysis that tracks the values stored
-in different variables in an effort to identify variables that are
-used in the same way in each submission.  They also show promising preliminary
-results for their approach.
-
-\subsection{Key Features:  Mining Hints from the AST}
-
-Our approach, which is very preliminary, ongoing research,  is also based on an analysis of
-the AST.
-Our hypothesis is that many novice-level exercises
-will have solutions that require certain structural
-{\em key features} in the code that we can detect automatically by
-comparing the AST of the student's code with the ASTs of known correct
-solutions.
-
-For example, Figures~\ref{fig:counta} and \ref{fig:countab} show two
-CloudCoder exercises that were assigned during the Winter 2013 term in
-CS-1 at 
-Small College \#1
-%Knox College.
-
-\begin{figure}[h]
-{\tt Write a static method called countA that takes a String as a
-parameter and returns the number of times that 'a' occurs in the String. }
-\caption{CountA:  An example CloudCoder exercise.}
-\label{fig:counta}
-\end{figure}
-
-\begin{figure}[h]
-{\tt Write a static method called countAB that takes a String as a
-parameter and returns the number of times that ab occurs (i.e. an 'a'
-immediately followed by 'b'). }
-\caption{CountAB: Another example CloudCoder exercise.}
-\label{fig:countab}
-\end{figure}
-
-First let us consider the {\tt CountA} exercise from
-Figure~\ref{fig:counta}.  As is typical in computer science, students have a lot of flexibility
-in how they solve this problem:  with a for loop or with a while loop, using the {\tt charAt()}
-method or the {\tt substring()} method or other String methods, and so on.  However, in
-practice, exercises are assigned in a context with other
-exercises, in-class examples, and course topics, which tends to limit
-the scope of possible student solutions, and to push students towards
-certain solutions and away from others.  For example, no student solved this
-exercise with a while loop, and all but one used {\tt charAt()}
-rather than other String operations.  In fact, most correct
-solutions looked something like Figure~\ref{fig:counta-solution}.
-
-\begin{figure}[h]
-\begin{lstlisting}
-public static int countAB(String str) {
-  int count=0;
-  for (int i=0; i<str.length();i++) {
-    if (str.charAt(i)=='a') {
-      count++;
-    }
-  }
-  return count;
-}
-\end{lstlisting}
-\caption{Typical solution for Exercise CountA depicted in Figure~\ref{fig:counta}.}
-\label{fig:counta-solution}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=90mm]{images/countA.pdf}
-\caption{Slightly simplified AST of the code in Figure~\ref{fig:counta-solution}}
-\label{fig:counta-ast}
-\end{figure}
-
-Compiling the method from Figure~\ref{fig:counta-solution} to an
-Abstract Syntax Tree (AST) produces a tree that looks something like
-Figure~\ref{fig:counta-ast} (the AST has been simplified slightly by
-collapsing nodes that are unimportant for this example).  We can identify
-several features of the AST that must be present to solve the
-exercise, such as a for loop, an if-statement inside the loop, and
-the use of a String method, typically {\tt charAt()}, to examine
-characters inside the loop.  We can also ensure that the components of
-the for loop are correct, i.e. that the loop counter starts at 0, that
-the loop goes up to the length of the String, and so on.  Assuming
-that student code compiles, all of these necessary features can be
-extracted from the AST.
-
-Now let's consider the {\tt CountAB} exercise depicted in
-Figure~\ref{fig:countab}.  Again, there were many correct solutions;
-however, many correct solutions looked something like
-Figure~\ref{fig:countab-solution}.
-
-\begin{figure}[h]
-\begin{lstlisting}
-public static int countAB(String str) {
-  int count=0;
-  for (int i=0; i<str.length()-1;i++) {
-    if (str.charAt(i)=='a' && 
-        str.charAt(i+1)=='b') {
-      count++;
-    }
-  }
-  return count;
-}
-\end{lstlisting}
-\caption{Typical solution for Exercise countAB depicted in Figure~\ref{fig:countab}.}
-\label{fig:countab-solution}
-\end{figure}
-
-Compiling the method in Figure~\ref{fig:countab-solution} into an
-abstract syntax tree (AST) produces a tree that looks like
-Figure~\ref{fig:countab-ast} (this result has been slightly simplified
-from the full output from the Eclipse JDT compiler).
-
-\begin{figure}[h]
-\includegraphics[width=90mm]{images/forloop-highlight.png}
-\caption{Slightly simplified AST of the code in Figure~\ref{fig:countab-solution}}
-\label{fig:countab-ast}
-\end{figure}
-
-Note the two circled sub-trees, or {\em features}, in the AST depicted
-in Figure~\ref{fig:countab-ast}.  Feature \#1, the {\em
-  length-1} feature, shows the loop iterating
-until the length of the String minus one.  The for loop requires a
-stop condition that is different
-than the loop stop condition used to solve the {\tt CountA} exercise from
-Figure~\ref{fig:counta}.  Crucially, once we know students are
-using a for loop to solve this exercise (which we can also
-extract from the AST), {\em we know exactly where in the AST to
-  look for the length-1 feature}.  If a submission has all the
-features except the {\em length-1} feature, we have a pretty good idea
-that the student hasn't figured out to change the stop condition of the
-for loop.  We could use the lack of features present in correct
-solutions to give us an idea of what things the student is missing.
-
-\subsection{Limits of Key Features Approach}
-
-Our approach of identifying ``key features'' has a number of
-weaknesses and limitations.  First, like other ITS hint-systems, it
-requires a reliable system for identifying the nearest previous successful
-submission to a student's current program state.  Work in this area is
-ongoing \cite{rivers-its2012,Jin:2012:PRA:2345840.2345889}.
-
-Second, features are based purely on the structure of the
-code, and are treated independently of one another, which is a
-simplification that throws out a lot of relevant contextual information.  For
-example, a student submission may have two for loops, one of which has
-an if statement nested inside it; our approach cannot determine which
-loop has the if statement inside it.
-
-Third, we do not perform any data-flow analysis (at least not yet), so
-we cannot identify if particular values are stored in variables.
-
-Next, we require compilable code, which can be a problem for the many
-students who struggle to get their code to compile
-\cite{Denny:2012:SEE:2325296.2325318}.
-
-Finally, while we focused mainly on exercises that required for loops
-to solve, we also assigned a number of exercises that
-use if statements and no loops.  For these exercises, we can figure
-out when students don't have 
-enough structure, such as when they don't have nested if statements
-for an exercise where most people who solves the exercise used nested if statements.  However,
-we cannot yet easily determine when students have the wrong boundary
-conditions in an if statement.
-
-Much more work is needed in this area.  We hope that 
-CloudCoder can help out by providing a rich data set for post-hoc
-studies, as well as a platform for testing out a hint-system.
+%
+%\section{Future Work: Towards an Intelligent Tutoring System}
+%
+%One of the long-term goals for CloudCoder is to use the data to
+%develop an Intelligent Tutoring System for programming assignments.
+%A web-based programming ITS could be used to support high school
+%AP courses in school districts that may lack computer science
+%teachers, or to help improve introductory CS courses by reducing
+%student frustration.
+%
+%To develop an ITS, we need a large supply of exercises; we hope
+%CloudCoder's permissible licensing will help on this front.  We also
+%need to provide students with ``hints'' when they are stuck.  To
+%provide hints, we can either write the hints ourselves, which is
+%labor-intensive, or generate them automatically by mining them out of
+%the dataset of prior submissions.  Obviously this second approach is
+%a much preferred method of creating hints.  However, identifying hints
+%automatically means that we need to match a student's current program
+%state with the most similar program state we've seen previously, 
+%figure out what successful next step the previous student took, and
+%then summarize this information into a hint.  This is a very
+%challenging problem.
+%
+%There are currently two promising lines of research
+%trying to address the issue of identifying the most similar previous submissions.  Rivers and Koedinger \cite{rivers-its2012} have developed a
+%system that computes a {\em canonicalized} version of each submission
+%by compiling to an Abstract Syntax Tree (AST) and then applying a
+%series of compiler transformations.  This work is ongoing, but
+%preliminary results are encouraging.
+%On the other hand, Jin et. al. \cite{Jin:2012:PRA:2345840.2345889, jin-kdd2011} perform
+%an inter-procedural data flow analysis that tracks the values stored
+%in different variables in an effort to identify variables that are
+%used in the same way in each submission.  They also show promising preliminary
+%results for their approach.
+%
+%\subsection{Key Features:  Mining Hints from the AST}
+%
+%Our approach, which is very preliminary, ongoing research,  is also based on an analysis of
+%the AST.
+%Our hypothesis is that many novice-level exercises
+%will have solutions that require certain structural
+%{\em key features} in the code that we can detect automatically by
+%comparing the AST of the student's code with the ASTs of known correct
+%solutions.
+%
+%For example, Figures~\ref{fig:counta} and \ref{fig:countab} show two
+%CloudCoder exercises that were assigned during the Winter 2013 term in
+%CS-1 at 
+%Small College \#1
+%%Knox College.
+%
+%\begin{figure}[h]
+%{\tt Write a static method called countA that takes a String as a
+%parameter and returns the number of times that 'a' occurs in the String. }
+%\caption{CountA:  An example CloudCoder exercise.}
+%\label{fig:counta}
+%\end{figure}
+%
+%
+%\begin{figure}[h]
+%{\tt Write a static method called countAB that takes a String as a
+%parameter and returns the number of times that ab occurs (i.e. an 'a'
+%immediately followed by 'b'). }
+%\caption{CountAB: Another example CloudCoder exercise.}
+%\label{fig:countab}
+%\end{figure}
+%
+%First let us consider the {\tt CountA} exercise from
+%Figure~\ref{fig:counta}.  As is typical in computer science, students have a lot of flexibility
+%in how they solve this problem:  with a for loop or with a while loop, using the {\tt charAt()}
+%method or the {\tt substring()} method or other String methods, and so on.  However, in
+%practice, exercises are assigned in a context with other
+%exercises, in-class examples, and course topics, which tends to limit
+%the scope of possible student solutions, and to push students towards
+%certain solutions and away from others.  For example, no student solved this
+%exercise with a while loop, and all but one used {\tt charAt()}
+%rather than other String operations.  In fact, most correct
+%solutions looked something like Figure~\ref{fig:counta-solution}.
+%
+%\begin{figure}[h]
+%\begin{lstlisting}
+%public static int countAB(String str) {
+%  int count=0;
+%  for (int i=0; i<str.length();i++) {
+%    if (str.charAt(i)=='a') {
+%      count++;
+%    }
+%  }
+%  return count;
+%}
+%\end{lstlisting}
+%\caption{Typical solution for Exercise CountA depicted in Figure~\ref{fig:counta}.}
+%\label{fig:counta-solution}
+%\end{figure}
+%
+%
+%\begin{figure}
+%\includegraphics[width=90mm]{images/countA.pdf}
+%\caption{Slightly simplified AST of the code in Figure~\ref{fig:counta-solution}}
+%\label{fig:counta-ast}
+%\end{figure}
+%
+%Compiling the method from Figure~\ref{fig:counta-solution} to an
+%Abstract Syntax Tree (AST) produces a tree that looks something like
+%Figure~\ref{fig:counta-ast} (the AST has been simplified slightly by
+%collapsing nodes that are unimportant for this example).  We can identify
+%several features of the AST that must be present to solve the
+%exercise, such as a for loop, an if-statement inside the loop, and
+%the use of a String method, typically {\tt charAt()}, to examine
+%characters inside the loop.  We can also ensure that the components of
+%the for loop are correct, i.e. that the loop counter starts at 0, that
+%the loop goes up to the length of the String, and so on.  Assuming
+%that student code compiles, all of these necessary features can be
+%extracted from the AST. 
+%
+%Now let's consider the {\tt CountAB} exercise depicted in
+%Figure~\ref{fig:countab}.  Again, there were many correct solutions;
+%however, many correct solutions looked something like
+%Figure~\ref{fig:countab-solution}.  
+%
+%\begin{figure}[h]
+%\begin{lstlisting}
+%public static int countAB(String str) {
+%  int count=0;
+%  for (int i=0; i<str.length()-1;i++) {
+%    if (str.charAt(i)=='a' && 
+%        str.charAt(i+1)=='b') {
+%      count++;
+%    }
+%  }
+%  return count;
+%}
+%\end{lstlisting}
+%\caption{Typical solution for Exercise countAB depicted in Figure~\ref{fig:countab}.}
+%\label{fig:countab-solution}
+%\end{figure}
+%
+%Compiling the method in Figure~\ref{fig:countab-solution} into an
+%abstract syntax tree (AST) produces a tree that looks like
+%Figure~\ref{fig:countab-ast} (this result has been slightly simplified
+%from the full output from the Eclipse JDT compiler).
+%
+%\begin{figure}[h]
+%\includegraphics[width=90mm]{images/forloop-highlight}
+%\caption{Slightly simplified AST of the code in Figure~\ref{fig:countab-solution}}
+%\label{fig:countab-ast}
+%\end{figure}
+%
+%Note the two circled sub-trees, or {\em features}, in the AST depicted
+%in Figure~\ref{fig:countab-ast}.  Feature \#1, the {\em
+%  length-1} feature, shows the loop iterating
+%until the length of the String minus one.  The for loop requires a
+%stop condition that is different
+%than the loop stop condition used to solve the {\tt CountA} exercise from
+%Figure~\ref{fig:counta}.  Crucially, once we know students are
+%using a for loop to solve this exercise (which we can also
+%extract from the AST), {\em we know exactly where in the AST to
+%  look for the length-1 feature}.  If a submission has all the
+%features except the {\em length-1} feature, we have a pretty good idea
+%that the student hasn't figured out to change the stop condition of the
+%for loop.  We could use the lack of features present in correct
+%solutions to give us an idea of what things the student is missing.
+%
+%
+%\subsection{Limits of Key Features Approach}
+%
+%Our approach of identifying ``key features'' has a number of
+%weaknesses and limitations.  First, like other ITS hint-systems, it
+%requires a reliable system for identifying the nearest previous successful
+%submission to a student's current program state.  Work in this area is
+%ongoing \cite{rivers-its2012,Jin:2012:PRA:2345840.2345889}.
+%
+%Second, features are based purely on the structure of the
+%code, and are treated independently of one another, which is a
+%simplification that throws out a lot of relevant contextual information.  For
+%example, a student submission may have two for loops, one of which has
+%an if statement nested inside it; our approach cannot determine which
+%loop has the if statement inside it.  
+%
+%Third, we do not perform any data-flow analysis (at least not yet), so
+%we cannot identify if particular values are stored in variables.
+%
+%Next, we require compilable code, which can be a problem for the many
+%students who struggle to get their code to compile
+%\cite{Denny:2012:SEE:2325296.2325318}.
+%
+%Finally, while we focused mainly on exercises that required for loops
+%to solve, we also assigned a number of exercises that
+%use if statements and no loops.  For these exercises, we can figure
+%out when students don't have 
+%enough structure, such as when they don't have nested if statements
+%for an exercise where most people who solves the exercise used nested if statements.  However,
+%we cannot yet easily determine when students have the wrong boundary
+%conditions in an if statement.
+%
+%Much more work is needed in this area.  We hope that 
+%CloudCoder can help out by providing a rich data set for post-hoc
+%studies, as well as a platform for testing out a hint-system.
 

### 2013-06-18 paper/paper.tex
change address
de-anonymize links
better example screenshot

@@
 David Hovemeyer\\
-       \affaddr{Kinsley Engineering Center}\\
+       \affaddr{Department of Physical Sciences}\\
        \affaddr{York College of Pennsylvania}\\
@@
 \centering
-\includegraphics[width=180mm]{images/screenshot}
-\caption{Screenshot of the main view in CloudCoder.}
+\includegraphics[width=5.5in]{images/screenshot4}
+\caption{Screenshot of the main view in CloudCoder}
 \label{screenshot}
@@
 CloudCoder is available under the GNU Affero open-source license.
-Binary distributions are available through our website (anonymized),
+Binary distributions are available through our website (\url{http://cloudcoder.org}),
 and the code is available on Github:
@@
 \vspace*{3mm} 
-%\url{https://github.com/cloudcoderdotorg/CloudCoder}
-(link anonymized)
+\url{https://github.com/cloudcoderdotorg/CloudCoder}
 \vspace*{3mm}
@@
 \vspace*{3mm} 
-% \url{https://cloudcoder.org/repo}
-(link anonymized)
+\url{https://cloudcoder.org/repo}
 \vspace*{3mm}

### 2013-06-19 paper/paper.tex
acknowedge special projects grant
tweak placement of fig.1
added a couple discussion questions

@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=5.5in]{images/screenshot4}
+\caption{Screenshot of the main view in CloudCoder}
+\label{screenshot}
+\end{figure*}
+
 \subsection{An Open Platform}
@@
 standard err to help students with debugging.
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[width=5.5in]{images/screenshot4}
-\caption{Screenshot of the main view in CloudCoder}
-\label{screenshot}
-\end{figure*}
 
@@
 
-\section{Discussion}
+\section{Discussion Questions}
 
-We could use community input into the CloudCoder project.  Specifically:
+%We could use community input into the CloudCoder project.  Specifically:
+%
+%\begin{itemize}
+%\item Who would be interested in adopting this system?  If you would
+%  consider trying out the system if we added a particular feature,
+%  what would that feature be?  We would like
+%  to expand our pilot group.
+%\item What studies could or should be done with this platform?  And
+%  would you like to be involved with these studies?
+%\item What studies can be done with the existing data we have
+%  collected?  We do not yet have well-developed tools for data
+%  analysis, but one data set has been anonymized and can be
+%  distributed to other researchers interested in our data.
+%\end{itemize}
 
-\begin{itemize}
-\item Who would be interested in adopting this system?  If you would
-  consider trying out the system if we added a particular feature,
-  what would that feature be?  We would like
-  to expand our pilot group.
-\item What studies could or should be done with this platform?  And
-  would you like to be involved with these studies?
-\item What studies can be done with the existing data we have
-  collected?  We do not yet have well-developed tools for data
-  analysis, but one data set has been anonymized and can be
-  distributed to other researchers interested in our data.
-\end{itemize}
+There are a number of questions, both philosophical and research-oriented,
+that we would like to address in the context of our future work with CloudCoder.
+
+{\bf Can the adoption of programming exercises lead to improved outcomes
+in introductory programming courses?}  This is the most obvious
+question.  One scenario we are particularly interested in is whether
+the use of programming exercises, when {\em added} to an existing course with
+no other changes in pedagogy, lead to improved outcomes.  We feel
+that this scenario is important because educators have limited time
+and resources available to change the way their courses are taught.
+If programming exercises are an easy change to implement, then
+if adopted on a wide scale they could have a significant impact.
+
+{\bf Can exercises be reused effectively outside of the course in which
+they were originally created?}  When assigning a programming exercise to
+students, the instructor can create one from scratch, or use a
+previously-written exercise.  Since creating good exercises is labor-intensive,
+being able to use a previously-written exercise represents a significant
+time savings.  However, finding an existing exercise that is {\em appropriate}
+is not necessarily a trivial task.  We are interested in finding
+out whether exercises written in the context of one course can be reused
+effectively in another.
 
@@
 
-Anonymized.
+Our work has been supported by a SIGCSE Special Projects grant (May 2012).
 

### 2013-06-19 paper/paper.tex
more discussion questions

@@
 \renewcommand{\floatpagefraction}{0.84}
+
+\newenvironment{denseItemize}{%
+\begin{list}{$\bullet$}{\setlength{\itemsep}{0in}\setlength{\parsep}{0in}\leftmargin=1em\labelwidth=1em}}{\end{list}}
 
@@
 out whether exercises written in the context of one course can be reused
-effectively in another.
+effectively in another.  There are many interesting aspects of
+this question:
+
+\begin{denseItemize}
+\item How can we help instructors search for appropriate exercises?
+\item How can we classify exercises?
+\item Can we allow CloudCoder users to {\em improve} existing exercises?
+\end{denseItemize}
+
+{\bf What guidance can we provide to instructors on which exercises to assign,
+and in what order?}  This question is an expanded form of the previous question:
+if a large repository of pre-written exercises is available,
+could we offer instructors sequences of exercises on a particular topic?
+
+{\bf How can programming exercises be integrated most effectively into a course?}
+Programming exercises can be used for many purposes: self-test questions
+to accompany readings and video lectures, in-class assessment (possibly
+in the context of peer instruction), extra practice for students who
+are struggling.  We would like to have specific, evidence-based recommendations
+that we can give to instructors who are considering using programming
+exercises in their courses.
 

### 2013-06-20 paper/paper.tex
philosophical questions
added a discussion question
more discussion questions from Jaime

@@
 
+{\bf Given an effective mechanism for sharing curriculum materials, could
+courses at a number of institutions converge towards an optimum?}
+Discuss.
+
+{\bf If exercises are shared freely, is the availability of solutions a concern?}
+Discuss.
+
+{\bf Should exercises be graded?}
+Discuss.
+
+{\bf What are the appropriate incentives to encourage students to attempt
+the exercises (and get the maximum benefit from them)?}
+Discuss.
+
 \section{Related Work}

### 2013-06-20 paper/paper.tex
another discussion question
a philosophical question

@@
 courses at a number of institutions converge towards an optimum?}
-Discuss.
+This question is more philosophical, and is based on the observation that
+curricula and pedagogy vary widely between instructors and institutions,
+even for the ``same'' course \cite{Hertz:2010:CCM:1734263.1734335}.  We find ourselves wondering
+if more effective ways ways of sharing curriculum materials would lead
+to the best materials and ideas propagating widely, and if high-quality introductory
+programming courses could created as a ``mash-up'' of materials created by
+many authors at many institutions.  CloudCoder and similar systems
+could be one mechanism enabling this type of transfer.
 
@@
 Discuss.
+
+{\bf What is the right balance between short programming exercises and
+  longer ``nifty'' assignments?}
+Discuss.
+
+{\bf What can we do to encourage students to use their time effectively outside of class?}
 

### 2013-06-20 paper/paper.tex
incentives

@@
 
-{\bf If exercises are shared freely, is the availability of solutions a concern?}
-Discuss.
+{\bf Should exercises be graded?
+If exercises are shared freely, is the availability of solutions a concern?}
+These questions are related: if exercises are available publicly,
+solutions will likely be available publicly as well.  (For example,
+many if not all CodingBat problems have solutions which can be found
+through a web search.)  If exercises are graded, then some students may
+choose to cheat rather than attempt to solve the problem honestly.
 
-{\bf Should exercises be graded?}
-Discuss.
-
-{\bf What are the appropriate incentives to encourage students to attempt
+{\bf What are appropriate incentives to encourage students to attempt
 the exercises (and get the maximum benefit from them)?}
-Discuss.
+If exercises are not graded, then what incentives can be used to
+encourage students to make a serious attempt at solving them?
 

### 2013-06-20 paper/paper.tex
get table 1 to float to the right place
more de-anonymizing

@@
 
+\begin{table*}
+\centering
+\begin{tabular}{| l | l | l | l |}
+\hline
+Institution & Course & Language & Enrollment\\
+\hline
+Canisius College & CS-1 & C/C++ & 30\\
+York College & CS-1 & C/C++ & 120\\
+Knox College \#1 & CS-1 & Java & 23\\
+University of Auckland & Programming for Engineers & C & 30\\
+\hline
+\end{tabular}
+\caption{CloudCoder pilot studies.}
+\label{tab:courses}
+\end{table*}
+
 CloudCoder is available under the GNU Affero open-source license.
@@
 
-\begin{table*}
-\centering
-\begin{tabular}{| l | l | l | l |}
-\hline
-Institution & Course & Language & Enrollment\\
-\hline
-Canisius College & CS-1 & C/C++ & 30\\
-York College & CS-1 & C/C++ & 120\\
-Knox College \#1 & CS-1 & Java & 23\\
-University of Auckland & Programming for Engineers & C & 30\\
-\hline
-\end{tabular}
-\caption{CloudCoder pilot studies.}
-\label{tab:courses}
-\end{table*}
-
-In the Winter 2013 term at Small College \#1, students taking CS-1 used {\tt CloudCoder}
+In the Winter 2013 term at Knox College, students taking CS-1 used {\tt CloudCoder}
 %
@@
 dataset from 120 students at 
-Small College \#2
-%York College 
+York College 
 once the Spring 2013 semester ends.  Our analysis of these datasets is

### 2013-06-20 paper/paper.tex
edited the abstract

@@
 We have built CloudCoder, an open platform for creating, assigning, and sharing
-short programming exercises for a variety of languages (we currently
-support C/C++, Java, Python and Ruby, and can add others).  Like other
-systems, CloudCoder is a web service that lets
-students write code directly in their web browser, click the
+short programming exercises for a variety of languages (currently C/C++, Java,
+Python and Ruby).  Like other similar systems, CloudCoder is web-based,
+letting students write code directly in a web browser, click the
 ``submit'' button, and receive immediate feedback.  Unlike other
@@
 
-In this work, we report on successful pilot studies of CloudCoder at several
-institutions.  We also describe on-going work to create a ``hint'' system
-for CloudCoder, similar to what one would expect from an Intelligent
-Tutoring System (ITS).
+In this paper, we report on successful pilot studies of CloudCoder at several
+institutions.
+% We also describe on-going work to create a ``hint'' system
+%for CloudCoder, similar to what one would expect from an Intelligent
+%Tutoring System (ITS).
+We also outline some research questions we hope to address in future work.
 
@@
 
-\section{Discussion Questions}
+\section{Questions}
 

### 2013-06-20 paper/paper.tex
tweak to abstract

@@
 
-We have built CloudCoder, an open platform for creating, assigning, and sharing
+We describe CloudCoder, an open platform for creating, assigning, and sharing
 short programming exercises for a variety of languages (currently C/C++, Java,
@@
 
-Our work has been supported by a SIGCSE Special Projects grant (May 2012).
+We would like to thank the anonymous reviewers for helpful suggestions
+for improving the paper.
+
+This work was supported by a SIGCSE Special Projects grant (May 2012).
 

### 2013-06-20 paper/paper.tex
various small tweaks

@@
 
-\keywords{CS1, novice programmers, data collection, student behavior,
-  intelligent tutoring system}
+\keywords{CS1, novice programmers, data collection, student behavior}
 
@@
 Recent years have seen an explosion of online services for learning
-to program:  Udacity \cite{udacity}, CodeAcademy \cite{codeacademy},
+to program:  Udacity \cite{udacity}, Codecademy \cite{codeacademy},
 ProgZoo \cite{progzoo}, 
@@
 Test-cases can be marked as ``secret'' so that students don't see their
-outputs to disincentivize students from writing a special case to
+outputs to discourage students from writing a special case to
 recognize each input.
@@
 
-There are a number of questions, both philosophical and research-oriented,
+There are a number of questions, both research-oriented and philosophical,
 that we would like to address in the context of our future work with CloudCoder.
@@
 the use of programming exercises, when {\em added} to an existing course with
-no other changes in pedagogy, lead to improved outcomes.  We feel
+no other changes in pedagogy, can result in improved outcomes.  We feel
 that this scenario is important because educators have limited time
@@
 out whether exercises written in the context of one course can be reused
-effectively in another.  There are many interesting aspects of
-this question:
+effectively in another.
+This question has several interesting aspects:
 

### 2013-06-20 paper/paper.tex
rewrite motivations for creating CloudCoder

@@
 improve introductory computer science.  Given all of these
-very good systems, one may ask why we thought it a good career move
-to build yet another system.  There are three main reasons:
+very good systems, one may ask why we thought
+%it a good career move
+%to build yet another system.
+the world needed another such system.
 
-%\begin{itemize}
+%There are three main reasons:
+%
+%%\begin{itemize}
+%
+%%\item 
+%First, we wanted a system that an instructor can easily integrate into an existing traditional course.
+%  Many massive open online courses (MOOCs), such those available through Udacity and
+%  Coursera, were not designed for a traditional classroom and cannot be easily incorporated into an existing course.  
+%  The infrastructure provided by a MOOC is geared towards the course
+%  taught through the MOOC, and is not really configurable or customizable.
+%
+% % systems, such as Udacity, Coursera, and CodeAcademy, are
+% %  massive open online courses (MOOCs) that cannot be easily
+% %  incorporated into a traditional classroom setting.  MOOCs can be wonderfully
+% %  enriching for students who take these courses.  However, because
+% %  MOOCs offer an entire
+% %  course, whatever online programming infrastructure they provide is part of that
+% %  course.
+%
+%%\item 
+%Second, we wanted an open, customizable system that focuses on short programming exercises that students
+%  can solve in their browser, and receive immediate feedback.  There are excellent systems that help
+%  faculty manage longer programming assignments
+%  \cite{Edwards:2008:WAG:1384271.1384371}, or homework and review questions
+%  \cite{turingscraft, Kumar:2005:GPA:1163405.1163408}.  Codingbat
+%  \cite{codingbat} is a fantastic free service that provides short
+%  programming exercises, but it is not easily customizable and does not provide
+%  access to all of the submissions submitted by students.
+%  We see CloudCoder as a complementary component to these resources.
+%
+%% Some existing systems, such as Problets \cite{Kumar:2005:GPA:1163405.1163408},  and Turingscraft
+%%   \cite{turingscraft}, provide homework exercises that can be incorporated
+%%   such as Codingbat \cite{codingbat},
+%%  , can be integrated into an existing course.
+%%   Codingbat, while free to use and fantastically useful, does not give faculty access to a very
+%%   rich data set
+%
+%%\item
+%Finally, we wanted exercises that could be freely modified and
+%redistributed.  Much of the excellent content available through MOOCs,
+%or the increasing amount of digital content available through textbook publishers, 
+%cannot be modified or re-used outside of those contexts.  This can be a significant
+%impediment to researchers trying to study the difficulty level of
+%various exercises, or to build intelligent tutoring systems (ITSs)
+%that require a large bank of questions and their solutions.
+%
+%%\end{itemize}
+%
+%% There are two reasons.  First, many of
+%% these systems, such are Udacity, Coursera, and CodeAcademy, are
+%% massive open online courses (MOOCs).  MOOCs typically provide
+%% supporting infrastructure that helps students learn to code in their
+%% browsers, and the course material offered by a MOOC can be a wonderfully enriching
+%% experience for someone who enrolls in the course.  However, it is not
+%% clear how faculty who teach more traditional courses can
+%% integrate any of the infrastructure provided by a MOOC into
+%% their curriculum.  Second, 
+%% the systems that could be easily integrated into an existing course, such as
+%% Codingbat or Turingscraft, are
+%% are either commercial, or closed, or both.  To our knowledge, none of
+%% these systems readily support faculty who want to 
+%% modify the code, host their own installation, extract data about
+%% student behaviors beyond the basic correctness of their submissions,
+%% or redistribute the programming exercises and assignments.
+%
+%We believe the academic community should have a fully open alternative
+%to the closed and/or commercial systems that have exploded onto the CS
+%Education scene.  Faculty can integrate CloudCoder into an existing course with minimal overhead,
+%can write their own content, use and modify content contributed by
+%others, customize the system, and collect data for educational studies.
 
-%\item 
-First, we wanted a system that an instructor can easily integrate into an existing traditional course.
-  Many massive open online courses (MOOCs), such those available through Udacity and
-  Coursera, were not designed for a traditional classroom and cannot be easily incorporated into an existing course.  
-  The infrastructure provided by a MOOC is geared towards the course
-  taught through the MOOC, and is not really configurable or customizable.
+Our primary motivation was {\em freedom}: we wanted a system where
+the software and exercises could be freely redistributed, and
+were not under the control of any central authority.  This motivation
+is not entirely philosophical: we felt that locking down either
+the software or the exercises would result in a barrier to adoption,
+limiting the potential impact of the system.
+As we will discuss in Section~\ref{sec:questions}, we think that there
+are some interesting research questions that can be addressed
+best through the existence of a ``free like air'' programming
+exercise system: for example, we are very interested in finding out
+whether exercises can be reused effectively in different courses
+at different institutions.
 
- % systems, such as Udacity, Coursera, and CodeAcademy, are
- %  massive open online courses (MOOCs) that cannot be easily
- %  incorporated into a traditional classroom setting.  MOOCs can be wonderfully
- %  enriching for students who take these courses.  However, because
- %  MOOCs offer an entire
- %  course, whatever online programming infrastructure they provide is part of that
- %  course.
+Another motivation was a desire to collect data on students' work
+on programming exercises as a way to study how students learn to program.
+For example, CloudCoder captures a fine-grained edit history of each
+student's work on each exercise.  Because the system is open source,
+we (and other researchers) can easily add other kinds of data collection
+to the system.
 
-%\item 
-Second, we wanted an open, customizable system that focuses on short programming exercises that students
-  can solve in their browser, and receive immediate feedback.  There are excellent systems that help
-  faculty manage longer programming assignments
-  \cite{Edwards:2008:WAG:1384271.1384371}, or homework and review questions
-  \cite{turingscraft, Kumar:2005:GPA:1163405.1163408}.  Codingbat
-  \cite{codingbat} is a fantastic free service that provides short
-  programming exercises, but it is not easily customizable and does not provide
-  access to all of the submissions submitted by students.
-  We see CloudCoder as a complementary component to these resources.
-
-% Some existing systems, such as Problets \cite{Kumar:2005:GPA:1163405.1163408},  and Turingscraft
-%   \cite{turingscraft}, provide homework exercises that can be incorporated
-%   such as Codingbat \cite{codingbat},
-%  , can be integrated into an existing course.
-%   Codingbat, while free to use and fantastically useful, does not give faculty access to a very
-%   rich data set
-
-%\item
-Finally, we wanted exercises that could be freely modified and
-redistributed.  Much of the excellent content available through MOOCs,
-or the increasing amount of digital content available through textbook publishers, 
-cannot be modified or re-used outside of those contexts.  This can be a significant
-impediment to researchers trying to study the difficulty level of
-various exercises, or to build intelligent tutoring systems (ITSs)
-that require a large bank of questions and their solutions.
-
-%\end{itemize}
-
-% There are two reasons.  First, many of
-% these systems, such are Udacity, Coursera, and CodeAcademy, are
-% massive open online courses (MOOCs).  MOOCs typically provide
-% supporting infrastructure that helps students learn to code in their
-% browsers, and the course material offered by a MOOC can be a wonderfully enriching
-% experience for someone who enrolls in the course.  However, it is not
-% clear how faculty who teach more traditional courses can
-% integrate any of the infrastructure provided by a MOOC into
-% their curriculum.  Second, 
-% the systems that could be easily integrated into an existing course, such as
-% Codingbat or Turingscraft, are
-% are either commercial, or closed, or both.  To our knowledge, none of
-% these systems readily support faculty who want to 
-% modify the code, host their own installation, extract data about
-% student behaviors beyond the basic correctness of their submissions,
-% or redistribute the programming exercises and assignments.
-
-We believe the academic community should have a fully open alternative
-to the closed and/or commercial systems that have exploded onto the CS
-Education scene.  Faculty can integrate CloudCoder into an existing course with minimal overhead,
-can write their own content, use and modify content contributed by
-others, customize the system, and collect data for educational studies.
+Finally, we wanted a system that supported a wide range of programming
+languages.  For example, relatively few programming exercise systems
+support C and C++.
 
@@
 
-\section{Questions}
+\section{Questions}\label{sec:questions}
 

### 2013-06-22 paper/paper.tex
some minor edits

@@
 
-Our primary motivation was {\em freedom}: we wanted a system where
+Our primary motivation in developing CloudCoder
+was {\em freedom}: we wanted a system where
 the software and exercises could be freely redistributed, and
@@
 \section{CloudCoder}
-We have designed CloudCoder to be an open platform for managing short
-programming exercises.  The CloudCoder platform has four design
-goals:  To be an open-platform for faculty to assign short programming
-exercises, to build a collection of redistributable programming
-exercises, to collect data about student behaviors as they solve
-exercises, and to explore ways of curating the collected data.
+%We have designed CloudCoder to be an open platform for managing short
+%programming exercises.  The CloudCoder platform has four design
+%goals:  To be an open-platform for faculty to assign short programming
+%exercises, to build a collection of redistributable programming
+%exercises, to collect data about student behaviors as they solve
+%exercises, and to explore ways of curating the collected data.
+
+In this section, we describe CloudCoder and discuss some initial experiences
+using it in introductory courses at several colleges and universities.
 
@@
 CloudCoder is an open platform for managing short programming
-exercises.  Each exercise can be solved with one method or one short
+exercises.
+Each exercise can be solved with one method or one short
 program, and could take only a few minutes to solve, or could require
@@
 programming assignments, such as those found at the {\em Nifty
-Assignments} session at SIGCSE.
+Assignments} session at SIGCSE \cite{Parlante:2013:NA:2445196.2445356}.
 
@@
 test cases, and feedback is provided directly in the browser.
-CloudCoder captures anything the code prints to standard out or
-standard err to help students with debugging.
+CloudCoder captures the code's standard output and
+standard error streams to help students with debugging.
 

### 2013-06-22 paper/paper.tex
various formatting and minor edits

@@
 Recent years have seen an explosion of online services for learning
-to program:  Udacity \cite{udacity}, Codecademy \cite{codeacademy},
-ProgZoo \cite{progzoo}, 
-Codingbat \cite{codingbat}, Coursera \cite{coursera}, Turingscraft
-\cite{turingscraft}, Problets \cite{Kumar:2005:GPA:1163405.1163408}, CS Circles \cite{Pritchard:2013:CCI:2445196.2445370} and
+to program:  Udacity\cite{udacity}, Codecademy\cite{codeacademy},
+ProgZoo\cite{progzoo}, 
+Codingbat\cite{codingbat}, Coursera\cite{coursera}, CodeLab\cite{turingscraft},
+Problets\cite{Kumar:2005:GPA:1163405.1163408}, CS Circles\cite{Pritchard:2013:CCI:2445196.2445370} and
 several others.  These systems are very powerful, and show great promise to
@@
 programming assignments, such as those found at the {\em Nifty
-Assignments} session at SIGCSE \cite{Parlante:2013:NA:2445196.2445356}.
+Assignments} session at SIGCSE\cite{Parlante:2013:NA:2445196.2445356}.
 
 Figure~\ref{screenshot} shows a screenshot of a CloudCoder exercise.
-The editor is derived from the Ace editor \cite{ace}, and supports
-syntax high-lighting for most programming languages, and
-auto-indentation.  
-A submit button runs the code against a series of
-test cases, and feedback is provided directly in the browser.
+CloudCoder uses the Ace editor\cite{ace}, which supports
+syntax highlighting and automatic indentation for most programming languages.
+The submit button runs the code against a series of
+test cases. Once the test results are received, they are presented
+directly in the browser, giving the student feedback on the
+correctness of the submission.
 CloudCoder captures the code's standard output and
@@
 currently support C/C++ less securely, although 
-recent work by Malan \cite{Malan:2013:CSS:2445196.2445242} on a
+recent work by Malan\cite{Malan:2013:CSS:2445196.2445242} on a
 ``sandbox'' for running untrusted student code can be leveraged to improve our security.
@@
 mechanism should be used to access the data.  Sanders
-et. al. \cite{Sanders:2008:DSE:1404520.1404534} reviewed a number of
+et. al.\cite{Sanders:2008:DSE:1404520.1404534} reviewed a number of
 the challenges involved in sharing empirical education data.
@@
 curricula and pedagogy vary widely between instructors and institutions,
-even for the ``same'' course \cite{Hertz:2010:CCM:1734263.1734335}.  We find ourselves wondering
+even for the ``same'' course\cite{Hertz:2010:CCM:1734263.1734335}.  We find ourselves wondering
 if more effective ways ways of sharing curriculum materials would lead
@@
 \cite{Jadud:2006:MTE:1151588.1151600}, along with follow-up work by
-others \cite{Norris:2008:CCQ:1384271.1384284} provides a starting point for
+others\cite{Norris:2008:CCQ:1384271.1384284} provides a starting point for
 work with program snapshots.
@@
 
-Lane et. al. describe Intention-Based Scoring (IBS) \cite{Lane:2005:ISA:1047344.1047471},
+Lane et. al. describe Intention-Based Scoring (IBS)\cite{Lane:2005:ISA:1047344.1047471},
 an approach to building an ITS for programming whereby human
@@
 The BlueJ team presents a compelling case for web-scale data
-collection \cite{Utting:2012:WDG:2361276.2361278}.  Their system is
+collection\cite{Utting:2012:WDG:2361276.2361278}.  Their system is
 more ambitious than CloudCoder, but the data collected by CloudCoder

### 2013-06-22 paper/paper.tex
cite AGPL
more small edits
editing exercises section, add some refs

@@
 Assignments} session at SIGCSE\cite{Parlante:2013:NA:2445196.2445356}.
+Instead, they are intended to be used as self-test exercises,
+practice problems, in-class lab activities, quiz problems, etc.
 
@@
 
-CloudCoder is available under the GNU Affero open-source license.
+CloudCoder is available under the GNU Affero General Public License\cite{AGPL}.
 Binary distributions are available through our website (\url{http://cloudcoder.org}),
 and the code is available on Github:
-
-\vspace*{3mm} 
+\begin{quote}
 \url{https://github.com/cloudcoderdotorg/CloudCoder}
-\vspace*{3mm}
+\end{quote}
 
@@
 
-Writing good homework exercises is difficult and
-time-consuming, and even the best textbooks are limited in the
-number and quality of practice exercises they can provide, to say
-nothing of a textbook's inability to provide timely feedback to
-students.
-Sharing exercises between faculty members is a much
-more efficient way to increase the number and quality of exercises at
-our disposal.
+%Writing good homework exercises is difficult and
+%time-consuming, and even the best textbooks are limited in the
+%number and quality of practice exercises they can provide, to say
+%nothing of a textbook's inability to provide timely feedback to
+%students.
+%Sharing exercises between faculty members is a much
+%more efficient way to increase the number and quality of exercises at
+%our disposal.
+Writing a good programming exercise is challenging.
+The instructions must be clear and concise.  The tests must
+include easy inputs as well as more difficult corner cases.
+If possible, tests should be constructed in a way that
+allows a partially-correct solution to pass at least some
+tests.  And, a good exercise should focus on a specific concept
+or language construct.
 
-Faculty who use CloudCoder can write their own exercises using our in-browser
-authoring tools, or can import exercises written by other faculty
-from a central repository at:
+%Faculty who use CloudCoder can write their own exercises using our in-browser
+%authoring tools, or can import exercises written by other faculty
+%from a central repository at:
 
-\vspace*{3mm} 
+Although instructors who use CloudCoder
+could choose to write from scratch all of the exercises they assign,
+our goal is to build a repository of good exercises
+written by many users at many institutions, made available under
+permissive licenses such as Creative Commons BY-SA\cite{CC-BY-SA}.
+Once published to the repository, an exercise is available
+for anyone who wishes to use it.
+The exercise repository is available at the URL
+\begin{quote}
 \url{https://cloudcoder.org/repo}
-\vspace*{3mm}
-
-Currently we have over 50 Java and 60 C/C++ exercises authored by 5
-different users.  By default,
-any new exercises written by faculty will be shared to the 
-central repository under a permissive Creative Commons license (BY-SA)
-that allows redistribution provided that the author is credited and
-redistribution is under the same license.  
-(Note that the repository shows the exercise description, but not the
-test cases or the solution, so an unscrupulous student cannot dig through the
-repository for solutions.  Importing exercises requires an account at
-the repository, which can be used to discourage this sort of
-undesirable behavior.).
+\end{quote}
+%
+At the time of writing (June 2013) the exercise repository contains about 50 Java and
+60 C/C++ exercises authored by 5 different users.  As more institutions
+start to use CloudCoder, we expect the number of exercises to increase.
+%By default,
+%any new exercises written by faculty will be shared to the 
+%central repository under a permissive Creative Commons license (BY-SA)
+%that allows redistribution provided that the author is credited and
+%redistribution is under the same license.  
+%(Note that the repository shows the exercise description, but not the
+%test cases or the solution, so an unscrupulous student cannot dig through the
+%repository for solutions.  Importing exercises requires an account at
+%the repository, which can be used to discourage this sort of
+%undesirable behavior.)
 
 CloudCoder supports two types of exercises:  {\em method} or {\em function-based}
-exercises, and {\em whole-program} exercises.  Function based exercises look
-a lot like Codingbat exercises:  Students are given a description of
-what the method should do, and a couple of example input/output pairs
-similar to a unit test.  Writing a new function-based exercise
-requires a description, some ``starter code'' (typically the method
+exercises, and {\em whole-program} exercises.  Function based exercises are
+very similar to CodingBat exercises: the student is required to write
+a function/method that takes some number of parameter values,
+does a computation, and returns a result value.
+Students are given a description of
+what the method should do, and a couple of example inputs with
+expected results.  Writing a new function-based exercise
+requires a description, some skeleton code (typically the method
 header, although exercises can provide as much or as little code as
-desired), and a series of inputs and their corresponding outputs.
-Test-cases can be marked as ``secret'' so that students don't see their
-outputs to discourage students from writing a special case to
+desired), and a series of inputs and their corresponding expected results.
+Test cases can be marked as ``secret'' so that students don't see their
+outputs, to discourage students from writing a special case to
 recognize each input.
 
-On the other hand, whole-program exercises
-have a description, but read inputs from
-standard-in and write to standard-out.  CloudCoder uses a
-simplified regular expression language to check student outputs for
-correctness.
+Whole-program exercises read input data from
+the standard input stream, perform a computation, and write output to the standard output stream.
+Expected output is checked by matching a regular expression against the
+program's standard output stream.
+Whole-program exercises are useful for two main reasons.  First,
+they can be assigned early in the course, before functions/methods
+have been introduced.  Second, they give the exercise author a
+way of checking expected output that is more powerful than matching a return
+value against a literal.  The main disadvantage is devising a
+regular expression that is precise enough to distinguish correct output
+from incorrect output, but forgiving enough to account for minor deviations
+in whitespace, capitalization, and formatting of output.
 
@@
 We run Java code relatively securely under a Security Manager, and 
-use Jython and JRuby to compile Python and Ruby to Java
+use Jython\cite{jython} and JRuby\cite{jruby} to compile Python and Ruby to Java
 bytecode, which we also run in a Security-Managed sandbox.  We
 currently support C/C++ less securely, although 
-recent work by Malan\cite{Malan:2013:CSS:2445196.2445242} on a
-``sandbox'' for running untrusted student code can be leveraged to improve our security.
+we hope to take advantage of a better sandboxing mechanism
+in the near future, possibly based on the work of
+David Malan\cite{Malan:2013:CSS:2445196.2445242} for Harvard's
+online CS50 course.
 

### 2013-06-22 paper/paper.tex
some revisions to building/testing

@@
 Ruby.  Building and testing takes place in a separate process which
-can be run on a separate machine for extra security, or on multiple
+can be run on a separate machine for extra security, and on multiple
 machines for scalability.
 
-We run Java code relatively securely under a Security Manager, and 
-use Jython\cite{jython} and JRuby\cite{jruby} to compile Python and Ruby to Java
-bytecode, which we also run in a Security-Managed sandbox.  We
-currently support C/C++ less securely, although 
+CloudCoder executes Java, Python, and Ruby code relatively securely under
+a Java Security Manager.
+(CloudCoder uses Jython\cite{jython} and JRuby\cite{jruby}
+to execute Python and Ruby submissions, allowing us to take advantage of
+the Java Security Manager for those languages.)
+CloudCoder's execution of C/C++ code is somewhat less secure, although 
 we hope to take advantage of a better sandboxing mechanism
@@
 David Malan\cite{Malan:2013:CSS:2445196.2445242} for Harvard's
-online CS50 course.
+online CS50 course.  However, because thus far the only users of
+CloudCoder have been instructors and students in introductory courses,
+the weak sandboxing has not been a significant concern.
 

### 2013-06-22 paper/paper.tex
a couple edits to data collection section

@@
 student's work on each exercise.  Because the system is open source,
-we (and other researchers) can easily add other kinds of data collection
-to the system.
+we (and other researchers) can easily add support for other kinds of data collection.
 
@@
 
-CloudCoder logs keystrokes in the online editor window, and 
-stores all submissions and test outcomes in a databases.
+As the student works on an exercises, CloudCoder logs all of the
+student's edits, and stores all submissions and test outcomes.
 We hope that the combination of exercises and data will give faculty
+and researchers
 insights into the difficulty level of different exercises, keep a
@@
 One question is how to manage the data collected.  
-Recent work on web-scale data collection by the BlueJ team
-\cite{Utting:2012:WDG:2361276.2361278} lays out a number of important
+Recent work on web-scale data collection by the
+BlueJ team\cite{Utting:2012:WDG:2361276.2361278} lays out a number of important
 questions about programming data, such as who should have access to the data, and what
 mechanism should be used to access the data.  Sanders
-et. al.\cite{Sanders:2008:DSE:1404520.1404534} reviewed a number of
+et.\ al.\cite{Sanders:2008:DSE:1404520.1404534} reviewed a number of
 the challenges involved in sharing empirical education data.

### 2013-06-22 paper/paper.tex
edits to hosting section

@@
 \subsection{Hosting CloudCoder}
+
 We recognize that ease of installation is a critical factor for
-adoption of a new pedagogical
-technology for busy
+adoption of a new pedagogical technology for busy
 faculty who may only have a few minutes to evaluate a new system.  We
@@
 
-Thus far each institution that has adopted CloudCoder has
-provided a Linux server that we configured for them.  This has made
-adoption very easy, but is unlikely to scale up as we acquire more
-users.  We have asked Amazon for Amazon Web Services credits to
-provide CloudCoder ``in the cloud'' to adopting institutions, and we
-hope to add support for Google App Engine this summer.  Ease of
-adoption is critical for CloudCoder.
+Setting up a CloudCoder instance requires two Linux servers: one to
+host the web application and database, and another to build and
+test the student submissions.  The installation process is reasonably well
+automated, and can be completed by someone comfortable with
+running Linux commands on remote servers in about an hour.
+We have received one report of a user successfully installing CloudCoder
+without assistance.  However, self-hosting is still a significant
+barrier to adoption, and we are investigating ways to make it
+easier for potential users to adopt CloudCoder.
+
+To date, each institution that has adopted CloudCoder in production has
+provided Linux servers that we configured for them.  This has made
+adoption very easy, but is not a model that is likely to scale well
+to larger numbers of institutions.  We recently received a grant
+from Amazon of credits for Amazon Web Services\cite{aws},
+which we plan to use to directly host CloudCoder for interested
+institutions.  We are also investigating easier ways for
+users to self-host, such as enabling CloudCoder to run on
+hosting services such as Google App Engine\cite{gae}.
+%  We have asked Amazon for Amazon Web Services credits to
+%provide CloudCoder ``in the cloud'' to adopting institutions, and we
+%hope to add support for Google App Engine this summer.  Ease of
+%adoption is critical for CloudCoder.
 

### 2013-06-22 paper/paper.tex
more acknowledgments before references
minor phrasing tweak

@@
 
-CloudCoder executes Java, Python, and Ruby code relatively securely under
+CloudCoder executes Java, Python, and Ruby code relatively securely using
 a Java Security Manager.
@@
 
+\section{Acknowledgements}
+
+We would like to thank the anonymous reviewers for helpful suggestions
+for improving the paper.
+
+This work was supported by a SIGCSE Special Projects grant (May 2012),
+a York College Summer Research Fellowship (Summer 2013), [TODO: Knox College
+grant], and [TODO: Amazon grant].
+
 \bibliographystyle{abbrv}
@@
 %Appendix A
-\section{Acknowledgements}
-
-We would like to thank the anonymous reviewers for helpful suggestions
-for improving the paper.
-
-This work was supported by a SIGCSE Special Projects grant (May 2012).
 

### 2013-06-22 paper/paper.tex
another minor phrasing tweak

@@
 automated, and can be completed by someone comfortable with
-running Linux commands on remote servers in about an hour.
+running Linux commands on remote servers in less than one hour.
 We have received one report of a user successfully installing CloudCoder

### 2013-06-22 paper/paper.tex
make pilot study table look nicer

@@
 \usepackage{hyperref}
+\usepackage{booktabs}
+
 \renewcommand{\topfraction}{0.9}
@@
 \centering
-\begin{tabular}{| l | l | l | l |}
-\hline
+\begin{tabular}{llll}
+\toprule
 Institution & Course & Language & Enrollment\\
-\hline
+\midrule
 Canisius College & CS-1 & C/C++ & 30\\
@@
 University of Auckland & Programming for Engineers & C & 30\\
-\hline
+\bottomrule
 \end{tabular}

### 2013-06-22 paper/paper.tex
fixes to pilot study table

@@
 \midrule
-Canisius College & CS-1 & C/C++ & 30\\
-York College & CS-1 & C/C++ & 120\\
-Knox College \#1 & CS-1 & Java & 23\\
+Canisius College & CS-1 & C++ & 30\\
+York College & CS-1 & C & 120\\
+Knox College & CS-1 & Java & 23\\
 University of Auckland & Programming for Engineers & C & 30\\

### 2013-06-22 paper/paper.tex
consistent formatting for citations

@@
 
-Jadud's pioneering early work on novice compilation behavior
-\cite{Jadud:2006:MTE:1151588.1151600}, along with follow-up work by
+Jadud's pioneering early work on novice compilation
+behavior\cite{Jadud:2006:MTE:1151588.1151600}, along with follow-up work by
 others\cite{Norris:2008:CCQ:1384271.1384284} provides a starting point for
@@
 
-Denny et. al. have studied novice syntax errors
-\cite{Denny:2012:SEE:2325296.2325318} using data collected from the CodeWrite system.
+Denny et. al. have studied novice syntax errors\cite{Denny:2012:SEE:2325296.2325318}
+using data collected from the CodeWrite system.
 CodeWrite is similar to CloudCoder, except that students create
@@
 
-Zingaro et. al. built the Python Classroom Response System
-\cite{Zingaro:2013:FCP:2445196.2445369}, a system similar to
+Zingaro et. al. built the Python Classroom Response
+System\cite{Zingaro:2013:FCP:2445196.2445369}, a system similar to
 CloudCoder, except that it focuses on supporting in-class programming
@@
 
-Pritchard et. al. present CS Circles
-\cite{Pritchard:2013:CCI:2445196.2445370}, an interactive on-line
+Pritchard et. al. present CS Circles\cite{Pritchard:2013:CCI:2445196.2445370},
+an interactive on-line
 Python curriculum that embeds short questions, some of which involve

### 2013-06-22 paper/paper.tex
another minor phrasing tweak

@@
 online CS50 course.  However, because thus far the only users of
-CloudCoder have been instructors and students in introductory courses,
+CloudCoder have been instructors and students,
 the weak sandboxing has not been a significant concern.

### 2013-06-23 paper/paper.tex
edits after a proofreading pass

@@
 is a completely open platform.  The code for the system is
-open-source, and exercises written for CloudCoder are shared to a
-central repository under a permissive Creative Commons license
-(BY-SA).  Finally, CloudCoder collects detailed data that faculty can use for
+open-source, and exercises written for CloudCoder may be shared to a
+central repository under permissive licenses such as Creative Commons BY-SA.
+Finally, CloudCoder collects detailed data that faculty can use for
 educational research.
@@
 Problets\cite{Kumar:2005:GPA:1163405.1163408}, CS Circles\cite{Pritchard:2013:CCI:2445196.2445370} and
-several others.  These systems are very powerful, and show great promise to
-improve introductory computer science.  Given all of these
-very good systems, one may ask why we thought
+many others.  These systems are powerful, and show great promise to
+improve introductory computer science.  Given the existence of all of these
+systems, one may ask why we thought
 %it a good career move
@@
 best through the existence of a ``free like air'' programming
-exercise system: for example, we are very interested in finding out
+exercise system: for example, we are interested in finding out
 whether exercises can be reused effectively in different courses
@@
 Figure~\ref{screenshot} shows a screenshot of a CloudCoder exercise.
-CloudCoder uses the Ace editor\cite{ace}, which supports
+CloudCoder uses Ace\cite{ace}, an open source web-based
+programmer's editor, which supports
 syntax highlighting and automatic indentation for most programming languages.
@@
 test cases. Once the test results are received, they are presented
-directly in the browser, giving the student feedback on the
+directly in the browser, giving the student immediate feedback on the
 correctness of the submission.
@@
 
-CloudCoder supports two types of exercises:  {\em method} or {\em function-based}
-exercises, and {\em whole-program} exercises.  Function based exercises are
+CloudCoder supports two types of exercises:  {\em function-based}
+exercises and {\em whole-program} exercises.  Function-based exercises are
 very similar to CodingBat exercises: the student is required to write
-a function/method that takes some number of parameter values,
+a function (or method, depending on the programming language) that takes some number of parameter values,
 does a computation, and returns a result value.
 Students are given a description of
-what the method should do, and a couple of example inputs with
+what the function should do, and a couple of example inputs with
 expected results.  Writing a new function-based exercise
@@
 Test cases can be marked as ``secret'' so that students don't see their
-outputs, to discourage students from writing a special case to
+inputs or expected results, to discourage students from writing a special case to
 recognize each input.
@@
 way of checking expected output that is more powerful than matching a return
-value against a literal.  The main disadvantage is devising a
+value against a literal.  The main disadvantage is
+the difficulty of devising a
 regular expression that is precise enough to distinguish correct output
-from incorrect output, but forgiving enough to account for minor deviations
+from incorrect output, but forgiving enough to allow minor deviations
 in whitespace, capitalization, and formatting of output.
@@
 CloudCoder's execution of C/C++ code is somewhat less secure, although 
-we hope to take advantage of a better sandboxing mechanism
-in the near future, possibly based on the work of
-David Malan\cite{Malan:2013:CSS:2445196.2445242} for Harvard's
-online CS50 course.  However, because thus far the only users of
+in the near future
+we hope to take advantage of a better sandboxing mechanism,
+possibly based on the work of
+David Malan for Harvard's online CS50 course\cite{Malan:2013:CSS:2445196.2445242}.
+However, because thus far the only users of
 CloudCoder have been instructors and students,
@@
 student's edits, and stores all submissions and test outcomes.
-We hope that the combination of exercises and data will give faculty
-and researchers
-insights into the difficulty level of different exercises, keep a
-record of common error patterns, and enable the development of
-innovative data analyses.
+Code edits are fine-grained---close to the level of keystroke
+logging.
+We hope that the data collected by CloudCoder will provide
+useful insights to faculty and researchers about
+exercises (their difficulty, which knowledge components they develop)
+and how students work on them (success patterns, error patterns, etc.)
+% combination of exercises and data will give faculty
+%and researchers
+%insights into the difficulty level of different exercises, keep a
+%record of common error patterns, and enable the development of
+%innovative data analyses.
 
-One question is how to manage the data collected.  
+One issue is how to manage the data collected.  
 Recent work on web-scale data collection by the
@@
 restrictions were put in place by the overseeing IRBs.  The Datashop also
-provides a suite of analysis and visualization tools for better making
+provides a suite of analysis and visualization tools for making
 sense of the ITS data.
@@
 faculty who may only have a few minutes to evaluate a new system.  We
-are committed to making the installation process as simple and painless as 
-possible.
+are committed to making the process
+of getting up and running as simple as possible.
+
+For evaluation purposes, we run a demo server
+(\url{https://cloudcoder.org/demo}) on which any instructor may request
+an account.  This provides an easy way for instructors to evaluate
+CloudCoder without installing the software.
 
@@
 to larger numbers of institutions.  We recently received a grant
-from Amazon of credits for Amazon Web Services\cite{aws},
+of credits for Amazon Web Services\cite{aws} from Amazon,
 which we plan to use to directly host CloudCoder for interested
 institutions.  We are also investigating easier ways for
-users to self-host, such as enabling CloudCoder to run on
+users to self-host, such as enabling CloudCoder to run on application
 hosting services such as Google App Engine\cite{gae}.
@@
 that we can give to instructors who are considering using programming
-exercises in their courses.
+exercises in their courses to guide them in using exercises in an effective manner.
 

### 2013-06-23 paper/paper.tex
a few more minor edits

@@
 
-We describe CloudCoder, an open platform for creating, assigning, and sharing
+In this paper, we describe CloudCoder, an open platform for creating, assigning, and sharing
 short programming exercises for a variety of languages (currently C/C++, Java,
@@
 
-In this paper, we report on successful pilot studies of CloudCoder at several
-institutions.
+We also report on successful pilot studies of CloudCoder at several
+institutions,
 % We also describe on-going work to create a ``hint'' system
@@
 %Tutoring System (ITS).
-We also outline some research questions we hope to address in future work.
+and outline research questions we hope to address in future work.
 
@@
 student's edits, and stores all submissions and test outcomes.
-Code edits are fine-grained---close to the level of keystroke
-logging.
+Code edits are fine-grained---close to the level of keystrokes.
 We hope that the data collected by CloudCoder will provide
 useful insights to faculty and researchers about
-exercises (their difficulty, which knowledge components they develop)
+exercises (their difficulty, which knowledge components they address)
 and how students work on them (success patterns, error patterns, etc.)

### 2013-06-24 paper/paper.tex
working on pilot studies section

@@
 \midrule
-Canisius College & CS-1 & C++ & 30\\
-York College & CS-1 & C & 120\\
-Knox College & CS-1 & Java & 23\\
+Canisius College & CS1 & C++ & 30\\
+York College & CS1 & C & 122\\
+Knox College & CS1 & Java & 23\\
 University of Auckland & Programming for Engineers & C & 30\\
@@
 academic year.  Table~\ref{tab:courses} shows a breakdown of the
-courses.  We tried to use CloudCoder at Auckland in Fall 2012 for a
-course with 700 students, but the free-tier of Amazon EC2 did not
-scale up well enough, so we only recorded partial data.  We have made
-significant improvements since then and plan to deploy CloudCoder for
+courses.  We also tried to use CloudCoder at the University of Auckland in Fall 2012 for a
+course with 700 students, but we encountered performance issues
+which prevented a full deployment.  We are working on scalability
+improvements, and we plan to deploy CloudCoder for
 the same large course in Fall 2013.
 
-In the Winter 2013 term at Knox College, students taking CS-1 used {\tt CloudCoder}
-%
-% TODO: \cite{Hovemeyer:2013:CBC:2445196.2445451}.  
-% 
+In the Winter 2013 term at Knox College, students taking CS1 used {\tt CloudCoder}.
 We collected over 6,000 submissions
-for 32 exercises by 23 students.  We expect to have a much larger
-dataset from 120 students at 
-York College 
-once the Spring 2013 semester ends.  Our analysis of these datasets is
-very preliminary at this point, but work is ongoing.
+for 32 exercises by 23 students. [TODO: what else?]
+
+In the Spring 2013 term at York College, we used CloudCoder in a CS1
+class, taught in C, with 122 students.  There were more than 56,000
+submissions for 68 exercises.  Because using CloudCoder was the
+only change in pedagogy compared to previous offerings of the course,
+we plan to analyze this data to see if we can find evidence that
+the exercises resulted in improved outcomes.
+
+%We expect to have a much larger
+%dataset from 120 students at 
+%York College 
+%once the Spring 2013 semester ends.  Our analysis of these datasets is
+%very preliminary at this point, but work is ongoing.
 
@@
 %CloudCoder exercises that were assigned during the Winter 2013 term in
-%CS-1 at 
+%CS1 at 
 %Small College \#1

### 2013-06-24 paper/paper.tex
some minor phrasing improvements

@@
 CloudCoder have been instructors and students,
-the weak sandboxing has not been a significant concern.
+the weak sandboxing of C/C++ submissions has not been a significant concern.
 
@@
 institutions.  We are also investigating easier ways for
-users to self-host, such as enabling CloudCoder to run on application
+users to self-host. One possibility is enabling CloudCoder to run on application
 hosting services such as Google App Engine\cite{gae}.

### 2013-06-24 paper/paper.tex
handle whitespace in et. al.
removed ITS from related work

@@
 We collected over 6,000 submissions
-for 32 exercises by 23 students. [TODO: what else?]
+for 32 exercises by 23 students.
 
@@
 
-Denny et. al. have studied novice syntax errors\cite{Denny:2012:SEE:2325296.2325318}
+Denny et.\ al. have studied novice syntax errors\cite{Denny:2012:SEE:2325296.2325318}
 using data collected from the CodeWrite system.
@@
 
-Lane et. al. describe Intention-Based Scoring (IBS)\cite{Lane:2005:ISA:1047344.1047471},
-an approach to building an ITS for programming whereby human
-judges to score or code {\em charettes}, or assignments students work on in a
-timed, closed-lab environment.  IBS derives from prior work on
-identifying ``bugs'' in online protocols, and as such attempts to
-identify places where students have a ``bug'' in their development
-process.  This approach, while labor-intensive, shows some promise in
-developing an ITS for programming assignments.
+%Lane et.\ al. describe Intention-Based Scoring (IBS)\cite{Lane:2005:ISA:1047344.1047471},
+%an approach to building an ITS for programming whereby human
+%judges to score or code {\em charettes}, or assignments students work on in a
+%timed, closed-lab environment.  IBS derives from prior work on
+%identifying ``bugs'' in online protocols, and as such attempts to
+%identify places where students have a ``bug'' in their development
+%process.  This approach, while labor-intensive, shows some promise in
+%developing an ITS for programming assignments.
 
-Zingaro et. al. built the Python Classroom Response
+Zingaro et.\ al. built the Python Classroom Response
 System\cite{Zingaro:2013:FCP:2445196.2445369}, a system similar to
@@
 
-Pritchard et. al. present CS Circles\cite{Pritchard:2013:CCI:2445196.2445370},
+Pritchard et.\ al. present CS Circles\cite{Pritchard:2013:CCI:2445196.2445370},
 an interactive on-line
@@
 %preliminary results are encouraging.
-%On the other hand, Jin et. al. \cite{Jin:2012:PRA:2345840.2345889, jin-kdd2011} perform
+%On the other hand, Jin et.\ al. \cite{Jin:2012:PRA:2345840.2345889, jin-kdd2011} perform
 %an inter-procedural data flow analysis that tracks the values stored

### 2013-06-24 paper/paper.tex
fix minor typo

--- 2013-06-24 paper/paper.tex
@@
 desired), and a series of inputs and their corresponding expected results.
-Test cases can be marked as ``secret'' so that students don't see their
+Test cases can be marked as ``secret'' so that students don't see the
 inputs or expected results, to discourage students from writing a special case to
@@
 
-Lane et. al. describe Intention-Based Scoring (IBS)\cite{Lane:2005:ISA:1047344.1047471},
+Lane et.\ al. describe Intention-Based Scoring (IBS)\cite{Lane:2005:ISA:1047344.1047471},
 an approach to building an ITS for programming whereby human
@@
 
-Zingaro et. al. built the Python Classroom Response
+Zingaro et.\ al. built the Python Classroom Response
 System\cite{Zingaro:2013:FCP:2445196.2445369}, a system similar to

### 2013-06-24 paper/paper.tex
avoid split infinitive

@@
 of credits for Amazon Web Services\cite{aws} from Amazon,
-which we plan to use to directly host CloudCoder for interested
+which we plan to use to host CloudCoder directly for interested
 institutions.  We are also investigating easier ways for

### 2013-06-24 paper/paper.tex
clarification about spring 2013 CS101 at YCP

@@
 submissions for 68 exercises.  Because using CloudCoder was the
-only change in pedagogy compared to previous offerings of the course,
+only major change in pedagogy compared to previous offerings of the course,
 we plan to analyze this data to see if we can find evidence that

### 2013-06-24 paper/paper.tex
clarification about screenshot
spelling corrections

@@
 ProgZoo\cite{progzoo}, 
-Codingbat\cite{codingbat}, Coursera\cite{coursera}, CodeLab\cite{turingscraft},
+CodingBat\cite{codingbat}, Coursera\cite{coursera}, CodeLab\cite{turingscraft},
 Problets\cite{Kumar:2005:GPA:1163405.1163408}, CS Circles\cite{Pritchard:2013:CCI:2445196.2445370} and
@@
 
-Figure~\ref{screenshot} shows a screenshot of a CloudCoder exercise.
+Figure~\ref{screenshot} shows a screenshot of the user interface
+students use to work on a CloudCoder exercise.
 CloudCoder uses Ace\cite{ace}, an open source web-based
@@
 Binary distributions are available through our website (\url{http://cloudcoder.org}),
-and the code is available on Github:
+and the code is available on GitHub:
 \begin{quote}
@@
 from incorrect output, but forgiving enough to allow minor deviations
-in whitespace, capitalization, and formatting of output.
+in white space, capitalization, and formatting of output.
 
@@
 
-\section{Acknowledgements}
+\section{Acknowledgments}
 

### 2013-06-25 paper/paper.tex
little more discussion, fixed acknowledgements

@@
   longer ``nifty'' assignments?}
-Discuss.
+There has been a lot of work in the CS Education community developing
+interesting and innovative introductory level programming assignments,
+such as the wonderful examples presented each year at the ``Nifty Assignments'' panel at
+SIGCSE.  Most introductory courses require at least a few
+longer nifty-style programming assignments, in addition to shorter
+programming homework assignments.  What is the proper balance between longer
+nifty-style assignments and shorter, more focused programming exercises?  Is 
+solving shorter programming exercises a pre-condition, or threshold skill, for
+completing assignments?
 
-{\bf What can we do to encourage students to use their time effectively outside of class?}
+{\bf Is the feedback provided by synchronous grading more effective
+  than feedback from asynchronous grading?}
+Faculty or Teaching Assistants who manually grade student programming
+assignments can give detailed and customized feedback to
+students, at the cost of some latency between the time the student
+submits the assignment and the time they receive their grade and the
+feedback.  Anecdotally, many faculty complain that students do not
+seem to pay much attention to this feedback, possibly because those
+points (marks) have already been lost and cannot be recovered.
+While automated systems cannot give feedback that is as detailed as manual
+grading, they can give feedback very quickly.
 
@@
 This work was supported by a SIGCSE Special Projects grant (May 2012),
-a York College Summer Research Fellowship (Summer 2013), [TODO: Knox College
-grant], and [TODO: Amazon grant].
+a York College Summer Research Fellowship (Summer 2013), an award from
+the Paul K Richter and Evalyn E. Cook Richter Memorial Trusts at Knox
+College, and an Amazon Web Services Teaching Grant.
 

### 2013-06-25 paper/paper.tex
editing to avoid a bad section break

@@
 
-Our primary motivation in developing CloudCoder
+Our primary motivation in developing CloudCoder,
+a new programming exercise system,
 was {\em freedom}: we wanted a system where
@@
 
-CloudCoder is available under the GNU Affero General Public License\cite{AGPL}.
+CloudCoder's source code is available under the GNU Affero General Public License\cite{AGPL}.
 Binary distributions are available through our website (\url{http://cloudcoder.org}),
@@
 very similar to CodingBat exercises: the student is required to write
-a function (or method, depending on the programming language) that takes some number of parameter values,
+a function (or method, depending on the programming language) that takes some parameter values,
 does a computation, and returns a result value.
@@
 a York College Summer Research Fellowship (Summer 2013), an award from
-the Paul K Richter and Evalyn E. Cook Richter Memorial Trusts at Knox
+the Paul K.\ Richter and Evalyn E.\ Cook Richter Memorial Trusts at Knox
 College, and an Amazon Web Services Teaching Grant.

### 2013-06-25 paper/paper.tex
get rid of window decorations in screenshot
minor formatting/consistency change
avoid a redundancy

@@
 \centering
-\includegraphics[width=5.5in]{images/screenshot4}
+\includegraphics[width=5.25in]{images/screenshot4}
 \caption{Screenshot of the main view in CloudCoder}
@@
 to larger numbers of institutions.  We recently received a grant
-of credits for Amazon Web Services\cite{aws} from Amazon,
+of credits for Amazon Web Services\cite{aws},
 which we plan to use to host CloudCoder directly for interested
@@
 
-In the Winter 2013 term at Knox College, students taking CS1 used {\tt CloudCoder}.
+In the Winter 2013 term at Knox College, students taking CS1 used CloudCoder.
 We collected over 6,000 submissions

### 2013-06-25 paper/paper.tex
minor edit

--- 2013-06-25 paper/paper.tex
@@
 ProgZoo\cite{progzoo}, 
-CodingBat\cite{codingbat}, Coursera\cite{coursera}, CodeLab\cite{turingscraft},
+Codingbat\cite{codingbat}, Coursera\cite{coursera}, CodeLab\cite{turingscraft},
 Problets\cite{Kumar:2005:GPA:1163405.1163408}, CS Circles\cite{Pritchard:2013:CCI:2445196.2445370} and
@@
 
-Figure~\ref{screenshot} shows a screenshot of the user interface
-students use to work on a CloudCoder exercise.
+Figure~\ref{screenshot} shows a screenshot of a CloudCoder exercise.
 CloudCoder uses Ace\cite{ace}, an open source web-based
@@
 Binary distributions are available through our website (\url{http://cloudcoder.org}),
-and the code is available on GitHub:
+and the code is available on Github:
 \begin{quote}
@@
 from incorrect output, but forgiving enough to allow minor deviations
-in white space, capitalization, and formatting of output.
+in whitespace, capitalization, and formatting of output.
 
@@
 to accompany readings and video lectures, in-class assessment (possibly
-in the context of peer instruction), extra practice for students who
+in the context of peer instruction), and extra practice for students who
 are struggling.  We would like to have specific, evidence-based recommendations
@@
   longer ``nifty'' assignments?}
-There has been a lot of work in the CS Education community developing
-interesting and innovative introductory level programming assignments,
-such as the wonderful examples presented each year at the ``Nifty Assignments'' panel at
-SIGCSE.  Most introductory courses require at least a few
-longer nifty-style programming assignments, in addition to shorter
-programming homework assignments.  What is the proper balance between longer
-nifty-style assignments and shorter, more focused programming exercises?  Is 
-solving shorter programming exercises a pre-condition, or threshold skill, for
-completing assignments?
+Discuss.
 
-{\bf Is the feedback provided by synchronous grading more effective
-  than feedback from asynchronous grading?}
-Faculty or Teaching Assistants who manually grade student programming
-assignments can give detailed and customized feedback to
-students, at the cost of some latency between the time the student
-submits the assignment and the time they receive their grade and the
-feedback.  Anecdotally, many faculty complain that students do not
-seem to pay much attention to this feedback, possibly because those
-points (marks) have already been lost and cannot be recovered.
-While automated systems cannot give feedback that is as detailed as manual
-grading, they can give feedback very quickly.
+{\bf What can we do to encourage students to use their time effectively outside of class?}
 
@@
 
-\section{Acknowledgments}
+\section{Acknowledgements}
 
@@
 This work was supported by a SIGCSE Special Projects grant (May 2012),
-a York College Summer Research Fellowship (Summer 2013), an award from
-the Paul K Richter and Evalyn E. Cook Richter Memorial Trusts at Knox
-College, and an Amazon Web Services Teaching Grant.
+a York College Summer Research Fellowship (Summer 2013), [TODO: Knox College
+grant], and [TODO: Amazon grant].
 

### 2013-06-25 paper/paper.tex
small change explaining motivation

@@
 were not under the control of any central authority.  This motivation
-is not entirely philosophical: we felt that locking down either
-the software or the exercises would result in a barrier to adoption,
+is not entirely philosophical: we felt that in any programming
+exercise system, locking down either
+software or exercises would result in a barrier to adoption,
 limiting the potential impact of the system.

### 2013-07-23 paper/talk.tex
started a beamer presentation

\documentclass{beamer}
\usepackage{graphicx}
\usetheme{Boadilla}

\begin{document}

\title[ICER 2013]{An Open System for Managing Short Programming Exercises}
\author[Papancea et.\ al.]{Andrei Papancea, Jaime Spacco, and David Hovemeyer${}^{\dagger}$}
\institute[Knox, YCP]{
\begin{minipage}[t]{1.25in}
\begin{center}
{{\ }\\
Knox College}
\vskip .1in
\includegraphics[height=.8in]{images/KnoxLogo}
\end{center}
\end{minipage}
\begin{minipage}[t]{1.25in}
\begin{center}
{York College\\
of Pennsylvania${}^{\dagger}$}
\vskip .1in
\includegraphics[height=.8in]{images/YCPLogo}
\end{center}
\end{minipage}
}
\date[August, 2013]{August 12, 2013}

\begin{frame}[plain]
  \titlepage
\end{frame}

\begin{frame}{Outline}

\begin{itemize}
  \item Background
  \item Description of CloudCoder
  \item Preliminary Experiences
  \item Questions
\end{itemize}

\end{frame}

\begin{frame}{Background}

\begin{itemize}
  \item Many students struggle in introductory programming courses
  \item Mastery of basic concepts and techniques is essential for success
  \item Short programming exercises: useful for
  \begin{itemize}
    \item Skills development
    \item Self-tests to accompany reading
    \item Extra practice
  \end{itemize}
\end{itemize}

\end{frame}

\end{document}

### 2013-07-23 paper/talk.tex
tweaks to talk title page, Makefile
some work on background slide
motivation slide

@@
 \institute[Knox, YCP]{
-\begin{minipage}[t]{1.25in}
+\begin{minipage}[t]{1.1in}
 \begin{center}
@@
 \end{minipage}
-\begin{minipage}[t]{1.25in}
+\begin{minipage}[t]{1.1in}
 \begin{center}
-{York College\\
-of Pennsylvania${}^{\dagger}$}
+{${}^{\dagger}$York College\\
+of Pennsylvania}
 \vskip .1in
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%% Outline %%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}{Outline}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%% Background %%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}{Background}
@@
   \item Mastery of basic concepts and techniques is essential for success
-  \item Short programming exercises: useful for
+  \item Short programming exercises:
+  \begin{itemize}
+    \item Students write or complete a short (5-15 line) program
+    \item Correctness is judged with a series of test cases
+    \item Student receives immediate feedback about results of tests
+  \end{itemize}
+  \item Exercises are useful for:
   \begin{itemize}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%% Why another system? %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{Why another system?}
+
+\begin{itemize}
+  \item There are lots of programming exercise systems already:
+  \begin{itemize}
+    \item Free: CodingBat, Codecademy
+    \item Commercial: CodeLab (TuringsCraft), MyProgrammingLab (Pearson)
+    \item {\em Do we really need another one?}
+  \end{itemize}
+  \item We did not know of one that
+  \begin{itemize}
+    \item Was open source
+    \item Encouraged users to create exercises
+    \item Allowed easy sharing of exercises
+    \item Supported a range of languages (including C/C++)
+  \end{itemize}
+\end{itemize}
+
+\end{frame}
+
 \end{document}

### 2013-07-27 paper/talk.tex
emph creating and sharing exercises
fleshed out Background and CloudCoder sections

@@
 
-\title[ICER 2013]{An Open System for Managing Short Programming Exercises}
+\title[ICER 2013]{An Open System for\\
+Managing Short Programming Exercises}
 \author[Papancea et.\ al.]{Andrei Papancea, Jaime Spacco, and David Hovemeyer${}^{\dagger}$}
@@
     \item Students write or complete a short (5-15 line) program
+    \item Interface is typically web-based
     \item Correctness is judged with a series of test cases
@@
   \begin{itemize}
-    \item Free: CodingBat, Codecademy
+    \item Free\footnote{As in beer}: CodingBat, Codecademy
     \item Commercial: CodeLab (TuringsCraft), MyProgrammingLab (Pearson)
@@
   \begin{itemize}
+    \item {\em Encouraged users to create and share exercises}
     \item Was open source
-    \item Encouraged users to create exercises
-    \item Allowed easy sharing of exercises
+    \item Collected rich data about how students work
     \item Supported a range of languages (including C/C++)
+  \end{itemize}
+  \item We created a new system, called CloudCoder, to fill this niche
+    \begin{itemize}
+    \item \url{http://cloudcoder.org}
     \end{itemize}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%% Freedom matters %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{Freedom matters}
+
+\begin{itemize}
+  \item The impact of a pedagogical technique or curriculum material
+        is a factor of how many students are exposed to it
+  \item Systematically eliminating barriers to adoption maximizes impact
+  \begin{itemize}
+    \item Requiring payment is a barrier to adoption
+    \item Restrictive licenses are a barrier to adoption
+    \item Control by a central authority is a barrier to adoption
+  \end{itemize}
+  \item Ideal situation: resources are available to anyone at any time for any purpose
+  \item Case in point...
+\end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%% Email #1 %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{Email \#1}
+
+\begin{center}
+\includegraphics[width=4.5in]{images/email1}
+\end{center}
+
+[I (DH) created an account and replied with the username and password.]
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%% Email #2 %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{Email \#2}
+
+\begin{center}
+\includegraphics[width=4.5in]{images/email2}
+\end{center}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%% Our philosophy %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{Our philosophy}
+
+\begin{itemize}
+  \item Make programming exercises written by many people available
+        for whatever uses are appropriate
+  \begin{itemize}
+    \item Use in courses
+    \item Use outside of courses? [i.e., the emails discussed previously]
+  \end{itemize}
+  \item Research how exercises are shared
+  \begin{itemize}
+    \item Find ways to make sharing more effective
+  \end{itemize}
+\end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%% CloudCoder %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{CloudCoder}
+
+\begin{itemize}
+  \item Open source (AGPL v3)
+  \item Supports exercises in Java, Python, Ruby, and C/C++
+  \item Collects fine-grained edit sequence of each student's work
+  \item Repository of freely-redistributable (CC-BY-SA) exercises:
+        \url{https://cloudcoder.org/repo}
+  \item Requires two Linux servers to host (one network-facing)
+    \begin{itemize}
+    \item Relatively easy to install
+    \item Known to support 60+ concurrent users with modest hardware
+    \item Load testing experiments successful with 350+ concurrent users
+          on higher-end hardware
+    \item Cost to host per course-semester: perhaps \$1-\$3 per student
+          (e.g., using Amazon EC2)
+    \end{itemize}
+\end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%% CloudCoder screenshot %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{CloudCoder screenshot}
+
+\begin{center}
+\vskip -.25in
+\includegraphics[width=4.5in]{images/screenshot4}
+\end{center}
+
+\end{frame}
+
 \end{document}

### 2013-07-29 paper/talk.tex
tweak placement of screenshot
phrasing tweak
better justification for free availability
take out emails for now - maybe add back later

@@
 Managing Short Programming Exercises}
-\author[Papancea et.\ al.]{Andrei Papancea, Jaime Spacco, and David Hovemeyer${}^{\dagger}$}
+\author[Papancea et.\ al.]{Andrei Papancea\textsuperscript{*}, Jaime Spacco\textsuperscript{*}, and David Hovemeyer\textsuperscript{$\dagger$}}
 \institute[Knox, YCP]{
@@
 {{\ }\\
-Knox College}
+\textsuperscript{*}Knox College}
 \vskip .1in
@@
 \begin{center}
-{${}^{\dagger}$York College\\
+{\textsuperscript{$\dagger$}York College\\
 of Pennsylvania}
@@
   \item The impact of a pedagogical technique or curriculum material
-        is a factor of how many students are exposed to it
+        depends on how many students are exposed to it
   \item Systematically eliminating barriers to adoption maximizes impact
@@
   \item Ideal situation: resources are available to anyone at any time for any purpose
-  \item Case in point...
+%  \item Case in point...
 \end{itemize}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%% Email #1 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{Email \#1}
-
-\begin{center}
-\includegraphics[width=4.5in]{images/email1}
-\end{center}
-
-[I (DH) created an account and replied with the username and password.]
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%% Email #2 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{Email \#2}
-
-\begin{center}
-\includegraphics[width=4.5in]{images/email2}
-\end{center}
 
@@
   \begin{itemize}
-    \item Use in courses
-    \item Use outside of courses? [i.e., the emails discussed previously]
+    \item Our need: exercises to assign in the context of a course
+    \item Related work: PCRS (Peterson et.\ al., SIGCSE 2013), in-class
+          exercises to support peer instruction
+    \item We can imagine lots of other uses: for example, Codecademy-style
+          self-learning
   \end{itemize}
-  \item Research how exercises are shared
+  \item Make the {\em sharing} of exercises a subject of research
   \begin{itemize}
-    \item Find ways to make sharing more effective
+    \item What kinds of exercises are most easily reused?
+    \item How can exercises be classified and organized?
   \end{itemize}
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%% Email #1 %%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{frame}{Email \#1}
+%
+%\begin{center}
+%\includegraphics[width=4.5in]{images/email1}
+%\end{center}
+%
+%[I (DH) created an account and replied with the username and password.]
+%
+%\end{frame}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%% Email #2 %%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{frame}{Email \#2}
+%
+%\begin{center}
+%\includegraphics[width=4.5in]{images/email2}
+%\end{center}
+%
+%\end{frame}
 
@@
 \begin{center}
-\vskip -.25in
+\vskip -.15in
 \includegraphics[width=4.5in]{images/screenshot4}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%% Experiences %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{Experiences}
+
+\end{frame}
+
 \end{document}

### 2013-08-08 paper/talk.tex
add conference name to title slide
spell Petersen correctly
revise estimate about number of users/hosting cost

@@
 }
-\date[August, 2013]{August 12, 2013}
+\date[August, 2013]{ICER\\
+August 12, 2013}
 
@@
     \item Our need: exercises to assign in the context of a course
-    \item Related work: PCRS (Peterson et.\ al., SIGCSE 2013), in-class
+    \item Related work: PCRS (Petersen et.\ al., SIGCSE 2013), in-class
           exercises to support peer instruction
@@
     \item Known to support 60+ concurrent users with modest hardware
-    \item Load testing experiments successful with 350+ concurrent users
-          on higher-end hardware
+    \item Load testing experiments successful with 700+ concurrent users
+          on moderate server hardware
+      \begin{itemize}
+      \item Compiling and testing submissions requires significant
+            CPU resources, but is easy to scale up by adding more
+            build servers
+      \end{itemize}
     \item Cost to host per course-semester: perhaps \$1-\$3 per student
@@
 
+\begin{itemize}
+\item Experiences:
+\end{itemize}
+
 \end{frame}

### 2013-08-08 paper/talk.tex
experiences slide
explanation of what a short exercise is

@@
   \end{itemize}
+\end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%% Short Exercises %%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}{Short Exercises}
+
+\begin{itemize}
+  \item Analogy: programming vs.\ learning to play a musical instrument
+  \item Larger assignment (like SIGCSE {\em Nifty Assignments}) :: Recital piece
+  \item Short exercise :: Playing a scale
 \end{itemize}
@@
 \begin{itemize}
+\item Pilot studies:
+  \begin{itemize}
+  \item Canisius College: 30 students, CS1, C++
+  \item York College: 122 students, CS1, C
+  \item Knox College: 23 students, CS1, Java
+  \item U.\ of Auckland: 30 students, CS1, C
+  \end{itemize}
 \item Experiences:
+  \begin{itemize}
+  \item The software works, students seem to like it
+  \item Writing exercises from scratch is hard
+  \item Repository has 100+ exercises, but not much reuse has occurred yet
+  \end{itemize}
+\item Can we scale it to larger courses?
+  \begin{itemize}
+\item September 2013: U.\ of Auckland: 700 students, CS1, C
+  \end{itemize}
 \end{itemize}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}{Questions}
+
+\end{frame}
+
 \end{document}

### 2013-08-08 paper/talk.tex
minor tweak
added questions

@@
 \begin{itemize}
-  \item Analogy: programming vs.\ learning to play a musical instrument
+  \item Analogy: learning programming vs.\ learning to play a musical instrument
   \item Larger assignment (like SIGCSE {\em Nifty Assignments}) :: Recital piece
@@
 
+\begin{itemize}
+\item Are short exercises useful?
+  \begin{itemize}
+  \item What experiments should we do to find out?
+  \end{itemize}
+\item How do we most effectively incorporate short exercises into a course?
+  \begin{itemize}
+  \item How do we know what exercises to use?
+  \item What mix of short exercises vs.\ larger assignments?
+  \end{itemize}
+\end{itemize}
+
 \end{frame}

### 2013-08-08 paper/talk.tex
tweaks to questions
hosting slide

@@
         \url{https://cloudcoder.org/repo}
+\end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%% CloudCoder screenshot %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{Hosting CloudCoder}
+
+\begin{itemize}
   \item Requires two Linux servers to host (one network-facing)
@@
           (e.g., using Amazon EC2)
+    \end{itemize}
+  \item {\bf We can host CloudCoder for your course!}
+    \begin{itemize}
+    \item We have a grant from Amazon to cover hosting costs for
+          a small number of institutions/courses
+    \item {\bf Talk to us if you are interested!}
     \end{itemize}
@@
   \begin{itemize}
-  \item What experiments should we do to find out?
+  \item Would {\em you} want to use short exercises in your course?
+  \item What experiments should we do to measure effect on outcomes?
   \end{itemize}
@@
   \begin{itemize}
-  \item How do we know what exercises to use?
+  \item How do we know what exercises should be assigned?
+  \item In what order should the exercises be assigned?
   \item What mix of short exercises vs.\ larger assignments?

### 2013-08-08 paper/talk.tex
move screenshot
added new screenshot
a few tweaks

@@
   \item Mastery of basic concepts and techniques is essential for success
-  \item Short programming exercises:
+  \item Short programming exercises ({\it {\` a} la} CodingBat):
   \begin{itemize}
@@
   \item Larger assignment (like SIGCSE {\em Nifty Assignments}) :: Recital piece
-  \item Short exercise :: Playing a scale
+  \item Short exercise :: Playing a scale, an {\' e}tude
 \end{itemize}
@@
   \end{itemize}
-  \item Ideal situation: resources are available to anyone at any time for any purpose
+  \item Ideal situation: resources are available to {\bf anyone} at {\bf any time} for {\bf any purpose}
 %  \item Case in point...
@@
   \item Open source (AGPL v3)
+  \item Web-based user interface
+    \begin{itemize}
+    \item 100\% HTML/CSS/JavaScript, no plugin required
+    \end{itemize}
   \item Supports exercises in Java, Python, Ruby, and C/C++
@@
 \end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%% CloudCoder screenshot %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{CloudCoder screenshot}
+
+\begin{center}
+\vskip -.15in
+\includegraphics[width=4.5in]{images/screenshot4}
+\end{center}
 
@@
     \item Relatively easy to install
-    \item Known to support 60+ concurrent users with modest hardware
-    \item Load testing experiments successful with 700+ concurrent users
+    \item Known to work: 60+ concurrent users with modest hardware
+    \item Load testing: with 700+ concurrent users
           on moderate server hardware
@@
 \end{itemize}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%% CloudCoder screenshot %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{CloudCoder screenshot}
-
-\begin{center}
-\vskip -.15in
-\includegraphics[width=4.5in]{images/screenshot4}
-\end{center}
 

### 2013-08-08 paper/talk.tex
more tweaks

@@
     \item {\em Encouraged users to create and share exercises}
+    \item Supported a range of languages (including C/C++)
     \item Was open source
     \item Collected rich data about how students work
-    \item Supported a range of languages (including C/C++)
   \end{itemize}
@@
   \item The impact of a pedagogical technique or curriculum material
-        depends on how many students are exposed to it
-  \item Systematically eliminating barriers to adoption maximizes impact
+        depends on how many students can access it
+  \item Systematically eliminating barriers to adoption and distribution maximizes impact
   \begin{itemize}
-    \item Requiring payment is a barrier to adoption
-    \item Restrictive licenses are a barrier to adoption
-    \item Control by a central authority is a barrier to adoption
+    \item Requiring payment is a barrier
+    \item Restrictive licenses are a barrier
+    \item Control by a central authority is a barrier
   \end{itemize}

### 2013-08-08 paper/talk.tex
a couple more small tweaks

@@
 \begin{itemize}
-  \item The impact of a pedagogical technique or curriculum material
+  \item The impact of a pedagogical technique or curriculum resource
         depends on how many students can access it
@@
       \begin{itemize}
-      \item Compiling and testing submissions requires significant
+      \item Compiling and testing submissions can require significant
             CPU resources, but is easy to scale up by adding more

### 2013-08-09 paper/talk.tex
emphasize difficulty of writing exercises
minor tweak
PCRS can import CloudCoder exercises!
current use?
authorship for PCRS paper

@@
     \item Correctness is judged with a series of test cases
-    \item Student receives immediate feedback about results of tests
+    \item Student receives immediate feedback
   \end{itemize}
@@
     \item Our need: exercises to assign in the context of a course
-    \item Related work: PCRS (Petersen et.\ al., SIGCSE 2013), in-class
+    \item Related work: PCRS (Zingaro et.\ al., SIGCSE 2013), in-class
           exercises to support peer instruction
+      \begin{itemize}
+      \item Summer 2013: PCRS can import exercises from the CloudCoder exercise repository!
+      \end{itemize}
     \item We can imagine lots of other uses: for example, Codecademy-style
@@
 \vskip -.15in
-\includegraphics[width=4.5in]{images/screenshot4}
+\includegraphics[width=4.5in]{images/screenshot5}
 \end{center}
@@
   \item The software works, students seem to like it
-  \item Writing exercises from scratch is hard
+  \item Writing exercises from scratch is hard!
   \item Repository has 100+ exercises, but not much reuse has occurred yet
@@
   \begin{itemize}
+  \item Do you use short exercises in your courses currently?
   \item Would {\em you} want to use short exercises in your course?

### 2013-08-09 paper/talk.tex
question about adoption
more work on questions slide

@@
   \item Do you use short exercises in your courses currently?
-  \item Would {\em you} want to use short exercises in your course?
+    \begin{itemize}
+    \item If so, how?
+    \item If not, would you be interested in starting to use them?
+    \end{itemize}
   \item What experiments should we do to measure effect on outcomes?
   \end{itemize}
-\item How do we most effectively incorporate short exercises into a course?
+\item How should we most effectively incorporate short exercises into a course?
   \begin{itemize}
@@
   \item What mix of short exercises vs.\ larger assignments?
+  \item How to motivate students to attempt exercises?
   \end{itemize}

### 2013-08-09 paper/talk.tex
grading questions?
tightening up to fit in 7 minutes

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%% Outline %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{Outline}
-
-\begin{itemize}
-  \item Background
-  \item Description of CloudCoder
-  \item Preliminary Experiences
-  \item Questions
-\end{itemize}
-
-\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%% Outline %%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{frame}{Outline}
+%
+%\begin{itemize}
+%  \item Background
+%  \item Description of CloudCoder
+%  \item Preliminary Experiences
+%  \item Questions
+%\end{itemize}
+%
+%\end{frame}
 
@@
   \item Mastery of basic concepts and techniques is essential for success
-  \item Short programming exercises ({\it {\` a} la} CodingBat):
+  \item Short programming exercises ({\` a} la CodingBat):
   \begin{itemize}
-    \item Students write or complete a short (5-15 line) program
-    \item Interface is typically web-based
+    \item Web-based UI
+    \item Students write or complete a short (5-15 line) function or program
     \item Correctness is judged with a series of test cases
@@
   \end{itemize}
-  \item Exercises are useful for:
+  \item Useful for:
   \begin{itemize}
@@
   \end{itemize}
-\end{itemize}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%% Short Exercises %%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}{Short Exercises}
-
-\begin{itemize}
-  \item Analogy: learning programming vs.\ learning to play a musical instrument
-  \item Larger assignment (like SIGCSE {\em Nifty Assignments}) :: Recital piece
-  \item Short exercise :: Playing a scale, an {\' e}tude
 \end{itemize}
@@
   \begin{itemize}
-    \item Free\footnote{As in beer}: CodingBat, Codecademy
+    \item Free: CodingBat, PracticeIt!, Codecademy
     \item Commercial: CodeLab (TuringsCraft), MyProgrammingLab (Pearson)
@@
   \begin{itemize}
-    \item {\em Encouraged users to create and share exercises}
+    \item Encouraged users to create and share {\bf freely-redistributable} exercises
     \item Supported a range of languages (including C/C++)
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%% CloudCoder %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{CloudCoder}
+
+\begin{itemize}
+  \item Open source (AGPL v3)
+  \item Supports exercises in Java, Python, Ruby, and C/C++
+  \item Web-based user interface
+    \begin{itemize}
+    \item 100\% HTML/CSS/JavaScript, no plugin required
+    \end{itemize}
+  \item Repository of freely-redistributable (CC-BY-SA) exercises:
+        \url{https://cloudcoder.org/repo}
+    \begin{itemize}
+    \item We encourage users to contribute, but do not require them to
+    \end{itemize}
+  \item Collects fine-grained edit sequence of each student's work
+\end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%% CloudCoder screenshot %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{CloudCoder screenshot}
+
+\begin{center}
+\vskip -.15in
+\includegraphics[width=4.5in]{images/screenshot5}
+\end{center}
+
+\end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%% Freedom matters %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{itemize}
-  \item The impact of a pedagogical technique or curriculum resource
-        depends on how many students can access it
-  \item Systematically eliminating barriers to adoption and distribution maximizes impact
+  %\item The impact of a pedagogical technique or curriculum resource
+  %      depends on how many students can access it
+  \item Goal: make programming exercises available to {\bf anyone} at {\bf any time} for {\bf any purpose}
   \begin{itemize}
-    \item Requiring payment is a barrier
-    \item Restrictive licenses are a barrier
-    \item Control by a central authority is a barrier
+    \item No payment
+    \item No restrictive licenses
+    \item No central control
   \end{itemize}
-  \item Ideal situation: resources are available to {\bf anyone} at {\bf any time} for {\bf any purpose}
-%  \item Case in point...
-\end{itemize}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%% Our philosophy %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{Our philosophy}
-
+  \item Exercises can be used in many contexts:
   \begin{itemize}
-  \item Make programming exercises written by many people available
-        for whatever uses are appropriate
-  \begin{itemize}
-    \item Our need: exercises to assign in the context of a course
-    \item Related work: PCRS (Zingaro et.\ al., SIGCSE 2013), in-class
-          exercises to support peer instruction
+    \item Our need: assign exercises in the context of a course
+    \item In-class exercises to support peer instruction
+          (PCRS, Zingaro et.\ al., SIGCSE 2013)
       \begin{itemize}
@@
       \end{itemize}
-    \item We can imagine lots of other uses: for example, Codecademy-style
-          self-learning
-  \end{itemize}
-  \item Make the {\em sharing} of exercises a subject of research
-  \begin{itemize}
-    \item What kinds of exercises are most easily reused?
-    \item How can exercises be classified and organized?
+    \item Codecademy-style self-learning?
   \end{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%% CloudCoder %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{CloudCoder}
-
-\begin{itemize}
-  \item Open source (AGPL v3)
-  \item Web-based user interface
-    \begin{itemize}
-    \item 100\% HTML/CSS/JavaScript, no plugin required
-    \end{itemize}
-  \item Supports exercises in Java, Python, Ruby, and C/C++
-  \item Collects fine-grained edit sequence of each student's work
-  \item Repository of freely-redistributable (CC-BY-SA) exercises:
-        \url{https://cloudcoder.org/repo}
-\end{itemize}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%% CloudCoder screenshot %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{CloudCoder screenshot}
-
-\begin{center}
-\vskip -.15in
-\includegraphics[width=4.5in]{images/screenshot5}
-\end{center}
-
-\end{frame}
-
 %%%%%%%%%%%%%%%%%%%%%%%% CloudCoder screenshot %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \item Known to work: 60+ concurrent users with modest hardware
-    \item Load testing: with 700+ concurrent users
+    \item Load testing: 700+ concurrent users
           on moderate server hardware
@@
       \item Compiling and testing submissions can require significant
-            CPU resources, but is easy to scale up by adding more
+            CPU resources, but are easy to scale up by adding more
             build servers
@@
   \end{itemize}
-\item How should we most effectively incorporate short exercises into a course?
+\item How to incorporate short exercises into a course most effectively?
   \begin{itemize}
@@
   \item How to motivate students to attempt exercises?
+    \begin{itemize}
+    \item Should they be graded?
+    \end{itemize}
   \end{itemize}

### 2013-08-12 paper/talk.tex
clarifications/simplifications
questions about classification/ranking

@@
     \item Encouraged users to create and share {\bf freely-redistributable} exercises
+    \item Was open source
     \item Supported a range of languages (including C/C++)
-    \item Was open source
     \item Collected rich data about how students work
@@
   \item Open source (AGPL v3)
-  \item Supports exercises in Java, Python, Ruby, and C/C++
-  \item Web-based user interface
-    \begin{itemize}
-    \item 100\% HTML/CSS/JavaScript, no plugin required
-    \end{itemize}
+  \item Supports Java, Python, Ruby, and C/C++
   \item Repository of freely-redistributable (CC-BY-SA) exercises:
@@
     \item We encourage users to contribute, but do not require them to
+    \end{itemize}
+  \item Web-based user interface
+    \begin{itemize}
+    \item 100\% HTML/CSS/JavaScript, no plugin required
     \end{itemize}
@@
           (PCRS, Zingaro et.\ al., SIGCSE 2013)
-      \begin{itemize}
-      \item Summer 2013: PCRS can import exercises from the CloudCoder exercise repository!
-      \end{itemize}
+      %\begin{itemize}
+      %\item Summer 2013: PCRS can import exercises from the CloudCoder exercise repository!
+      %\end{itemize}
     \item Codecademy-style self-learning?
@@
   \end{itemize}
+\item How do we classify exercises?
+\item How do we rank exercises?
 \item How to incorporate short exercises into a course most effectively?
   \begin{itemize}
-  \item How do we know what exercises should be assigned?
-  \item In what order should the exercises be assigned?
+  \item What exercises should be assigned? In what order?
   \item What mix of short exercises vs.\ larger assignments?
-  \item How to motivate students to attempt exercises?
+  \item How to motivate students to attempt them?
     \begin{itemize}
